### 2023-09-14 paper/main.tex
add initial latex files

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\title{Title Goes Here}

\begin{document}
\maketitle

\begin{abstract}
  Abstract goes here.
\end{abstract}

\section{Introduction}

hello world

\end{document}

### 2023-09-15 paper/main.tex
get most of the text into latex, but it's not rendering correctly...

@@
 
-\usepackage{graphicx}
-\usepackage{xspace}
+\usepackage{algorithm,algorithmic}  %% typset algorithms
 \usepackage{amssymb}
@@
 \usepackage{amsthm}
+\usepackage{graphicx}
+%%\usepackage{newpxtext}  %% for _e.g._ macro.
+\usepackage{textgreek}
+\usepackage{xspace}
+
+% use these if you also want π to work in math mode
+\usepackage{newunicodechar}
+\newunicodechar{Π}{\ifmmode\Pi\else\textPi\fi}
 
@@
 
+\author{First1 Last1}
+%%\authornote{with author1 note}        %% and a great more deleted...
+
 \begin{document}
+
+\newtheorem{definition}{Definition}
+
+\ExplSyntaxOn  %% for e.g. macro
+
+\newcommand { \eg }  %% https://tex.stackexchange.com/a/639374/244649
+{ \textit{e.g.}
+  \peek_meaning_ignore_spaces:NTF .
+    { \skip_horizontal:n { -.3ex } \use_none:n }
+    { \peek_meaning_ignore_spaces:NF , { \skip_horizontal:n { -.3ex } } }
+}
+\newcommand { \ie }
+{ \textit{i.e.}
+  \peek_meaning_ignore_spaces:NTF .
+    { \skip_horizontal:n { -.3ex } \use_none:n }
+    { \peek_meaning_ignore_spaces:NF , { \skip_horizontal:n { -.3ex } } }
+}
+
+\newcommand{\termOfArt}[1]{\textbf{#1}}
+
 \maketitle
@@
 
-hello world
+\dots  %% TODO: intorduction per se?
+
+\subsection{Motivation}
+
+  As in all kinds of engineering; in cryptography it is critical to know that the thing you are building actually does what you intend.
+  In cryptography however, the intended behavior is often expressed as a negation,
+  \eg secret information should \textit{never} be derivable from such'n'such surface of the system.
+  Observing of the system's behavior provides limited evidence that the system is actually secure in the intended ways.
+  Formal proofs of security are a popular and well-suited solution to this problem, however:
+
+\begin{itemize}
+    \item As consumers, we are concerned that a great deal of cyber-security software has been and is being produced without formal verification.
+    \item As engineers, we would like to not waste time attempting to prove the security of systems which, unbeknownst to us, are not actually secure.
+    \item As product managers, we would like to have sanity checks
+          that complicated software systems actually have the security properties their constituent source-code claims to be proven to have.
+\end{itemize}
+
+In the work we particularly focus on Secure Multi-Party Computation (MPC),
+a cryptographic primitive with wide-ranging applications but frustrating performance limitations.
+Substantial work has going into improving the performance of MPC systems [citation].
+Additionally, in some cases performance can be improved by using a bespoke protocol instead of a general-purpose MPC system [do we have any evidence of this?].
+We accelerate this work by providing a tool, comparable to a property-based test,
+for conveniently spot-checking the security of an MPC implementation.
+
+\subsection{Contributions}
+
+In this work we present and evaluate a statistical test of the security of MPC protocols.
+First we show how MPC security relates to inference and equivalence of probabilistic distributions,
+and then we demonstrate that existing tests of distributional equivalence can, with minor modification, be used to refute MPC security.
+Our tool is adaptable to the wide range of cryptographic security properties that can be defined in terms of distributional equivalence,
+and our tool is applicable to any software system that
+A) can be run 10k-1M times without great inconvenience, and
+B) allows certain intermediate values (of which consumed randomness is the least convenient) to be captured for analysis.
+
+\section{Background}
+
+\subsection{MPC security}
+
+Secure Multi-Party Computation (MPC or SMPC, depending on the text)
+is a cryptographic process that allows some group of participating parties or machines
+to compute a function that depends on all of their private inputs,
+without any of them learning anything except the final result.
+Implementations that work on arbitrary functions (representable as circuit computations in some field)
+have existed since [...],
+but their use by the public is limited by their performance.
+Some performance limitations are fundamental to the problem
+(roughly, every branch of the computation must be executed so that the control-flow does not itself leak information),
+but techniques have been found to improve performance somewhat.
+
+The traditional expression of MPC assumes a pre-determined function F, "the functionality",
+mapping the tuple of all participants' input vectors to the respective tuple of output vectors
+(or to distributions of outputs, if F is non-deterministic).
+In an \termOfArt{ideal world}, a trusted third party (the \termOfArt{ideal functionality}) would accept input messages
+from all other parties, compute F, and then send the parties their respective outputs.
+A \termOfArt{real world} protocol Π is said to correctly implement F if, for all possible inputs,
+Π results in all parties getting the same outputs as they would have in the ideal world.
+Defining the security of Π requires a clear threat model, and requires defining the participants' "views".
+
+\termOfArt{Passive Security} supposes a computationally-bounded adversary
+that can see everything happening inside a fixed subset of the participating machines.
+This adversary cannot cause the parties to deviate from the protocol Π in any way;
+we are interested in whether it can learn anything beyond the outputs of the "corrupted" machines
+by observing the execution of Π.
+As an example, this notion of security might be of interest if Alice trusted Bob to participate
+in a distributed computation with her, but did not trust him to properly store or destroy the log-files
+generated.
+The \termOfArt{ideal world view} of a party is what it sees in the ideal world: just its inputs and outputs.
+The \termOfArt{real world view} of a party is everything it sees in the real world while executing its part of Π:
+its inputs, outputs, all messages it receives, and all the bits of random tape it uses.
+
+\begin{definition}[Passive MPC Security]
+A protocol Π,
+involving participating parties P,
+that correctly implements functionality F,
+is secure against a passive adversary controlling C$\subset$P iff
+there exists a poly-time-computable function SIM s.t.
+....
+\end{definition}
+
+This definition represents security by a transitive argument:
+Since the adversary in the ideal world can compute all the data the real-world adversary starts with,
+anything the real-world adversary can compute the ideal-world adversary can also compute!
+
+In practice, security is usually quantified over all possible corruptions C
+where |C|$\leq$|P|/x for some x, but this is not fundamental and we only consider a single C at a time.
+
+Definition 1 is tight, in the sense that attempting to relax it invariable either
+allows a "secure" protocol to do something bad, or requires introducing restrictions
+on the kinds of functionality to which the definition may be applied.
+(Of course the idea that a secure protocol \textit{can't} do anything bad implies an assumption that computing
+the functionality is a good thing. [move to footnote])
+That said, some simplification is possible for our purposes:
+Since the corrupted parties are assumed to follow Π faithfully, we are fine to limit ourselves to
+choreographic protocols in which all messages sent are expected and all messages expected are received.
+In this context, and because the only source of non-determinism is the parties' random tapes
+(which are captured in the real-world views),
+we can safely ignore any messages sent \textit{within} C.
+Then we can collapse all the corrupted parties C into a single party C,
+and similarly all the other parties into a single party H (for "honest").
+Finally, in this work we assume that protocols implement their functionalities \textit{correctly},
+and usually leave the corresponding F implicit.
+In other words, we define F to be whatever mapping results from the execution of Π.
+This present work describes a system for property-based testing of MPC \textit{security};
+a corresponding test of \textit{correctness} is trivial for any pre-defined F.
+
+\termOfArt{Active Security} (sometimes "malicious security")
+supposes a computationally bounded adversary that can cause corrupted machines to take arbitrary actions,
+breaking the choreography,
+forging messages (to such extent as that's actually possible),
+lying about computation results,
+and/or playing along with the protocol for any period of time.
+In contrast to passive adversaries, which are weaker than the threat models most real programmers worry about,
+an active adversary subsumes any threat model (including adaptive corruption, depending how it's framed).
+The corresponding security definition is a bit more involved, and we omit it for brevity.
+Since we're not directly testing for active security in this work,
+all else that needs to be said about it is that
+any protocol which is not \termOfArt{passive secure} is automatically not \termOfArt{active secure}.
+
+[Other aspects we may wish to cover:
+ vacuous protocols,
+ vacuous security,
+ infertabilty of functionality,
+ comparison to control flow properties,
+ review of other work,
+ review of stuff we tried that didn't work...
+ ]
+
+[ computer proofs of MPC paper; (double-check malicions case...)
+  l-obliv
+  wysteria et al
+  probabilistic separation logic]
+
+\subsection{Inference in PPLs}
+
+\subsection{Property-based testing}
+
+\section{Related Work}
+
+\subsection{Verification of MPC}
+
+[Computer-aided proofs for multiparty computation with active security](https://arxiv.org/pdf/1806.07197.pdf)
+
+\subsection{Languages built inside of MPC implementations}
+
+\subsection{PPLs for Inference}
+
+dice
+
+\subsection{PPL proof assistants}
+
+lilac
+
+\subsection{Probabilistic tests of distributional equality}
+
+[decision trees paper]()
+
+\section{The tool in theory}
+
+... Position the equivalence in the definition of MPC security...
+
+To actually test a protocol, one uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
+One of the two models has access to the "real world" views, and the only has access to the "ideal world" views.
+In principal, \textit{any} ML system is ok to use; the choice is partly an engineering question
+and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
+Generating the data for these tests requires running the protocol many times.
+The process is enumerated in Algorithm 1.
+
+Note that the data must be understood as vectors/matricies/tensors of bits. [is this actually important? are there exceptions/precondidtions?]
+
+\begin{algorithm}  %% https://tex.stackexchange.com/a/219820/244649
+  \caption{Does it need a name?}
+  \begin{algorithmic}[1]
+  \STATE Set a chosen number of iterations `iters`,
+         number of rows of data on which to train each model `trainN`,
+         and number of rows of data on which to test each model `testN`.
+  \STATE Choose $\alpha$, your probability of randomly flagging a secure protocol as insecure.
+  \FOR {`i <- [1..iters]`, calculate scores for the real and ideal models:}
+    \STATE Run the protocol `trainN` times, collecting the corrupt views and honest secrets.
+    \STATE Train the first ML model on the collected data.
+           The real-world views are the input features, and the honest secrets are the labels. [I think this is generic ML terminology?]
+    \STATE Train the second ML model the same way, but occlude (omit or overwrite) view data to only reveal the ideal-world view.
+    \STATE Run the protocol `testN` more times, again collecting the relevant data.
+    \STATE The score for each model is the $L^1$ distance between the testing labels and the respective model outputs
+           (\ie the number of bits wrong when run on the `testN` rows of test data).
+  \ENDFOR
+  \STATE Using the `iters` samples of "real-world" and "ideal-world" scores,
+         calculate the $p$-value for the hypothesis that the distribution of real-world scores is better on average than the ideal-world distribution.
+  \IF {$p \leq \alpha$} \RETURN `INSECURE` \ELSE \RETURN `SECURE` \ENDIF
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Issue: false positives}
+
+What our tool actually invalidates is that the views are uniformly random within the slice compatible with the obseved inputs and outputs.
+
+\section{The tool in practice}
+
+[How weedy do we want to get here?]
+
+\subsection{A Choreography Language}
+
+Our python script for actually performing the test consumes a stream of fixed-length bit vectors.
+(In practice, this is a CSV. Annotation of which columns are available in the real and ideal worlds is provided in the header.)
+[Does our tool strictly need the vectors to be fixed width? Could someone do this kind of test on a system with branching control flow?]
+To facilitate analysis of the tool, we implemented a choreographic programming language, CHO,
+with surrounding tooling to support large batches of parallel evaluations of a single program and piping of the generated data.
+The design of CHO is driven by the contextual assumptions of MPC security.
+[Do we want a BNF? Semantics?]
+In addition to bit-level computations and transmission of single bits between arbitrary parties,
+each party has its own read-only tape for secret inputs and random bits
+(CHO is also a probabilistic language),
+and its own tape for output bits.
+CHO supports macros, but they are extremely limited.
+Finally, CHO supports 1-of-N oblivious transfer as a primitive operation.
+This allows us to test (correct and incorrect) implementations of the classic GMW protocol.
+
+\subsection{Random generation of programs}
+
+The goal of this work is a tool that can detect mistakes in MPC protocols.
+Such a tools is only useful if there if new protocols and implementations are being written.
+Therefore, while it is critical to demonstrate that our tool can detect errors manually edite into known-secure protocols,
+we \textit{also} need to show that the tool can detect a lack of security in entirely novel and maximally diverse protocols.
+We attempt to show this by randomly generating hundreds [can we say thousands?] of protocols and testing them with different test parameters.
+
+The probability that a randomly generated protocol will be MPC secure is not completely negligible.
+As an easy example, if no communicated values ever depend on secret inputs, then the protocol meets the definition of security!
+(Never mind that such a protocol is useless; \textit{none} of the protocols we generated are \textit{useful}.)
+That said, for reasonable generator settings we find that \textit{very few} generated protocols
+continue to appear secure as we increase the power of the test.
+
+[Probably a lot of the details of the generator go in the appendix?]
+
+\subsection{Performance analysis}
+
+\section{Future Work}
+
+\section{Conclusions}
 

### 2023-09-17 python/beaver_circuit_generator.py
add command-line arguments to circuit generators

@@ -1,13 +1,26 @@
 from circuit_generator import gensym, emit, gen_circuit, gen_randomness
 import random
+from argparse import (ArgumentParser, FileType)
+
+argp = ArgumentParser(description="Run the decision tree test on a csv of data.")
+argp.add_argument("input_file", type=FileType('r', 1, encoding='utf_8', errors='strict'))
+argp.add_argument("output_file", type=FileType('w', 1, encoding='utf_8', errors='strict'))
+argp.add_argument("--bias_sharing", action="store", type=int, default=0,
+                  help="Bias randomness used in secret sharing")
+argp.add_argument("--bias_and", action="store", type=int, default=0,
+                  help="Bias randomness used for AND gates")
+argp.add_argument("--accidental_secret", action="store", type=float, default=0.0,
+                  help="Rate of accidentally sending secret inputs to corrupt party")
+argp.add_argument("--accidental_gate", action="store", type=float, default=0.0,
+                  help="Rate of accidentally sending shares of gate outputs to corrupt party")
+args = argp.parse_args()
 
 config = {
-    'bias_sharing': 0,
-    'bias_and': 0,
-    'accidental_secret': 0.0,
-    'accidental_gate': 1.0,
+    'bias_sharing': args.bias_sharing,
+    'bias_and': args.bias_and,
+    'accidental_secret': args.accidental_secret,
+    'accidental_gate': args.accidental_gate,
     }
-
 
 def gen_xor(a, b):
     out = gensym('g')
@@ -98,4 +111,4 @@ MACRO reveal(P1(x1), P2(x2)) AS
 ENDMACRO
 """
 
-gen_circuit(config, generators, header)
+gen_circuit(config, generators, header, args.input_file, args.output_file)

### 2023-09-17 python/circuit_generator.py

@@ -18,18 +18,12 @@ global_f = None
 def emit(s=''):
     global_f.write(s + '\n')
 
-def gen_circuit(config, generators, header):
+def gen_circuit(config, generators, header, circuit_file, output_file):
     global global_f
-    assert len(sys.argv) == 3
 
-    circuit_filename = sys.argv[1]
-    output_filename = sys.argv[2]
+    global_f = output_file
 
-    global_f = open(output_filename, "w")
-
-    with open(circuit_filename, "r") as c:
-        circuit_lines = c.readlines()
-
+    circuit_lines = circuit_file.readlines()
 
     num_gates, num_wires = [int(x) for x in circuit_lines[0].strip().split(' ')]
     inputs_line = circuit_lines[1].strip().split(' ')

### 2023-09-17 python/gmw_circuit_generator.py

@@ -1,13 +1,26 @@
 from circuit_generator import gensym, emit, gen_circuit, gen_randomness
 import random
+from argparse import (ArgumentParser, FileType)
+
+argp = ArgumentParser(description="Run the decision tree test on a csv of data.")
+argp.add_argument("input_file", type=FileType('r', 1, encoding='utf_8', errors='strict'))
+argp.add_argument("output_file", type=FileType('w', 1, encoding='utf_8', errors='strict'))
+argp.add_argument("--bias_sharing", action="store", type=int, default=0,
+                  help="Bias randomness used in secret sharing")
+argp.add_argument("--bias_and", action="store", type=int, default=0,
+                  help="Bias randomness used for AND gates")
+argp.add_argument("--accidental_secret", action="store", type=float, default=0.0,
+                  help="Rate of accidentally sending secret inputs to corrupt party")
+argp.add_argument("--accidental_gate", action="store", type=float, default=0.0,
+                  help="Rate of accidentally sending shares of gate outputs to corrupt party")
+args = argp.parse_args()
 
 config = {
-    'bias_sharing': 0,
-    'bias_and': 0,
-    'accidental_secret': 0.0,
-    'accidental_gate': 1.0,
+    'bias_sharing': args.bias_sharing,
+    'bias_and': args.bias_and,
+    'accidental_secret': args.accidental_secret,
+    'accidental_gate': args.accidental_gate,
     }
-
 
 def gen_xor(a, b):
     out = gensym('g')
@@ -62,4 +75,4 @@ MACRO reveal(P1(x1), P2(x2)) AS
 ENDMACRO
 """
 
-gen_circuit(config, generators, header)
+gen_circuit(config, generators, header, args.input_file, args.output_file)

### 2023-09-17 python/circuit_experiment.py
update circuit experiment file to perform breakages

@@ -5,8 +5,21 @@ from datetime import datetime
 
 now = datetime.now()
 
+def gen_circuit_file(circuit_file, cho_file, protocol_type, config):
+    bias_sharing, bias_and, accidental_secret, accidental_gate = config
 
-def run_experiment(cho_file, iters, train, test, results_filename):
+    circuit_filename = 'circuits/' + circuit_file
+
+    p1 = subprocess.Popen(['python', f'python/{protocol_type}_circuit_generator.py',
+                           circuit_filename, cho_file,
+                           '--bias_sharing', str(bias_sharing),
+                           '--bias_and', str(bias_and),
+                           '--accidental_secret', str(accidental_secret),
+                           '--accidental_gate', str(accidental_gate)])
+    p1.wait()
+
+def run_experiment(cho_filename, iters, train, test, results_filename,
+                   protocol_type, config, circuit_name):
     # generate data
     data_csv = open('data.csv', "w")
     stderr_log = open('stderr.log', "w")
@@ -14,8 +27,6 @@ def run_experiment(cho_file, iters, train, test, results_filename):
     iters = str(iters)
     train = str(train)
     test = str(test)
-
-    cho_filename = 'circuits/' + cho_file
 
     p1 = subprocess.Popen(['/usr/bin/time', '-f', '\'%e\'', 'cabal', 'exec', 'd-tree-data', '--',
                            iters, train, test, '-f', cho_filename],
@@ -47,8 +58,9 @@ def run_experiment(cho_file, iters, train, test, results_filename):
     with open('stderr.log', "r") as f:
         d_tree_time = float(f.read().replace('\'', ''))
 
+    config_str = ','.join([str(x) for x in config])
     with open(results_filename, 'a') as f:
-        line = f'{cho_file},{iters},{train},{test},{pval},{data_time},{d_tree_time}\n'
+        line = f'{circuit_name},{protocol_type},{iters},{train},{test},{pval},{data_time},{d_tree_time},{config_str}\n'
         print(line)
         f.write(line)
 
@@ -56,37 +68,45 @@ def run_experiment(cho_file, iters, train, test, results_filename):
 iters = [64, 128]
 trains = [2**i for i in range(7, 11)]
 
-cho_names = [
[…]

### 2023-09-17 python/circuit_experiment.py
make the training range smaller

@@ -66,7 +66,7 @@ def run_experiment(cho_filename, iters, train, test, results_filename,
 
 
 iters = [64, 128]
-trains = [2**i for i in range(7, 11)]
+trains = [2**i for i in range(7, 10)]
 
 circuit_names = [
     'adder64.txt',

### 2023-09-18 python/circuit_experiment.py
attepmting to make single-threaded

@@ -1,5 +1,6 @@
 import os
 import subprocess
+import sys
 
 from datetime import datetime
 
@@ -69,14 +70,16 @@ iters = [64, 128]
 trains = [2**i for i in range(7, 10)]
 
 circuit_names = [
-    'adder64.txt',
-    'mult64.txt',
-    'aes_128.txt',
-    'divide64.txt',
-    'FP-add.txt',
-    'FP-mul.txt',
-    'sha256.txt',
-    'FP-div.txt']
+        cn
+        for cn in ['adder64.txt',
+                   'mult64.txt',
+                   'aes_128.txt',
+                   'divide64.txt',
+                   'FP-add.txt',
+                   'FP-mul.txt',
+                   'sha256.txt',
+                   'FP-div.txt']
+        if cn in sys.argv]
 
 protocol_types = ['gmw', 'beaver']
 

### 2023-09-19 python/circuit_experiment.py
revert singlethreading

@@ -70,16 +70,14 @@ iters = [64, 128]
 trains = [2**i for i in range(7, 10)]
 
 circuit_names = [
-        cn
-        for cn in ['adder64.txt',
-                   'mult64.txt',
-                   'aes_128.txt',
-                   'divide64.txt',
-                   'FP-add.txt',
-                   'FP-mul.txt',
-                   'sha256.txt',
-                   'FP-div.txt']
-        if cn in sys.argv]
+    'adder64.txt',
+    'mult64.txt',
+    'aes_128.txt',
+    'divide64.txt',
+    'FP-add.txt',
+    'FP-mul.txt',
+    'sha256.txt',
+    'FP-div.txt']
 
 protocol_types = ['gmw', 'beaver']
 

### 2023-09-19 python/d-tree-csv.py
support multiple input files to d-tree-csv

@@ -13,7 +13,8 @@ from sys import (exit, stderr)
 from warnings import catch_warnings, filterwarnings
 
 argp = ArgumentParser(description="Run the decision tree test on a csv of data.")
-argp.add_argument("file", type=FileType('r', 1, encoding='utf_8', errors='strict'))
+#argp.add_argument("file", type=FileType('r', 1, encoding='utf_8', errors='strict'))
+argp.add_argument('file', type=FileType('r', 1, encoding='utf_8', errors='strict'), nargs='+')
 argp.add_argument("iterations", type=int)
 argp.add_argument("trainingN", type=int)
 argp.add_argument("testingN", type=int)
@@ -33,7 +34,7 @@ except FileNotFoundError:
 except ValueError:
     exit(f"{__file__} was unable to parse \"./cores\" as an integer.")
 
-df = pd.read_csv(args.file)
+df = pd.concat([pd.read_csv(f) for f in args.file])
 
 if len(df) != NUM_ITERS * PER_ITER:
     print(f"Using {PER_ITER} rows for {NUM_ITERS} iterations, but there are {len(df)} rows!",

### 2023-09-19 python/circuit_experiment.py
modify circuit experiment to use multiple cores for generating data

@@ -1,10 +1,21 @@
 import os
 import subprocess
 import sys
+import math
 
 from datetime import datetime
 
 now = datetime.now()
+
+try:
+    with open("./cores") as cores_file:
+        CORES = int(cores_file.read())
+except FileNotFoundError:
+    exit(f"{__file__} was unable to find a file named \"./cores\" to tell it how many machine cores to use.")
+except ValueError:
+    exit(f"{__file__} was unable to parse \"./cores\" as an integer.")
+
+assert math.log(CORES, 2).is_integer()
 
 def gen_circuit_file(circuit_file, cho_file, protocol_type, config):
     bias_sharing, bias_and, accidental_secret, accidental_gate = config
@@ -22,31 +33,39 @@ def gen_circuit_file(circuit_file, cho_file, protocol_type, config):
 def run_experiment(cho_filename, iters, train, test, results_filename,
                    protocol_type, config, circuit_name):
     # generate data
-    data_csv = open('data.csv', "w")
-    stderr_log = open('stderr.log', "w")
-
+    iters_per_core = int(iters) // CORES
+    iters_per_core = str(iters_per_core)
     iters = str(iters)
     train = str(train)
     test = str(test)
 
-    p1 = subprocess.Popen(['/usr/bin/time', '-f', '\'%e\'', 'cabal', 'exec', 'd-tree-data', '--',
-                           iters, train, test, '-f', cho_filename],
-                          stdout=data_csv, stderr=stderr_log)
-    p1.wait()
+    pids = []
+    for i in range(CORES):
+        data_csv = open(f'data{i}.csv', "w")
+        stderr_log = open(f'stderr{i}.log', "w")
 
-    stderr_log.close()
-    data_csv.close()
+        p1 = subprocess.Popen(['/usr/bin/time', '-f', '\'%e\'', 'cabal', 'exec', 'd-tree-data', '--',
+                            iters_per_core, train, test, '-f', cho_filename],
+                            stdout=data_csv, stderr=stderr_log)
+        pids.append((p1, data_csv, stderr_log))
 
-    with open('stderr.log', "r") as f:
+    for pid, data_csv, stderr_log in pids:
+        pid.wait()
+        stderr_log.close()
+        data_csv.close()
+
+    with open('stderr0.log', "r") as f:
         data_time = float(f.read().replace('\'', ''))
 
     # run d-trees
     stdout_log = open('stdout.log', "w")
[…]

### 2023-09-19 python/circuit_experiment.py
capture all the times taken for data-generation threads

@@ -54,8 +54,11 @@ def run_experiment(cho_filename, iters, train, test, results_filename,
         stderr_log.close()
         data_csv.close()
 
-    with open('stderr0.log', "r") as f:
-        data_time = float(f.read().replace('\'', ''))
+    def retrieve_time(i):
+        with open(f'stderr{i}.log', "r") as f:
+            return float(f.read().replace('\'', ''))
+
+    data_time = max(retrieve_time(i) for i in range(CORES))
 
     # run d-trees
     stdout_log = open('stdout.log', "w")

### 2023-09-26 paper/main.tex
wip
wrap up old laptop
sketch evaluation

@@
 
-\ExplSyntaxOn  %% for e.g. macro
-
-\newcommand { \eg }  %% https://tex.stackexchange.com/a/639374/244649
+\ExplSyntaxOn    %% https://tex.stackexchange.com/a/639374/244649
+\newcommand { \eg }
 { \textit{e.g.}
@@
 }
+\ExplSyntaxOff
 
@@
 
+\section{Evaluation}
+
+\subsection{Secure Protocols Evaluated}
+
+\begin{table}
+\centering
+  \begin{tabular}{|l | l|l|}
+    \hline
+    \textbf{Protocol} & \textbf{Purpose} & \textbf{Ref.} \\
+    \hline
+    $n$-bit Addition & Custom & TODO \\
+    Oblivious Transfer & Building block & TODO \\
+    OT Extension & Building block & TODO \\
+    Beaver Triple Generation & Building block & TODO \\
+    GMW & Circuit evaluation & TODO \\
+    MPC w/ Beaver Triples & Circuit evaluation & TODO \\
+    \hline
+  \end{tabular}
+  \caption{Secure protocols used in our evaluation.}
+\end{table}
+
+\begin{table}
+\centering
+  \begin{tabular}{|l |l|l|}
+    \hline
+    \textbf{Circuit} & \textbf{AND Gates} & \textbf{Total Gates} \\
+    \hline
+    64-bit adder & 63 & 376 \\
+    64-bit multiplier & 4,033 & 13,675 \\
+    AES-128 & 6,400 & 36,663 \\
+    SHA-256 & 22,573 & 135,073 \\
+    Floating-point division & 82,269 & 184,007 \\
+    \hline
+  \end{tabular}
+  \caption{Circuits used in our evaluation.}
+\end{table}
+
+\subsection{Experiment 1: Scalability for Secure Protocols}
+
+\subsection{Experiment 2: Bug-Finding for Insecure Protocol Mutations}
+
+\subsection{Experiment 3: Scalability and Bug-Finding for Random Programs}
+
 \section{Future Work}

### 2023-09-26 paper/main.tex
some protocol descriptions

@@
 \newunicodechar{Π}{\ifmmode\Pi\else\textPi\fi}
+
+\renewcommand{\paragraph}[1]{\vspace*{2pt}\noindent\textbf{#1}}
+
+\newcommand{\system}{\textsc{SystemName}\xspace}
 
@@
 
-\section{Evaluation}
+\section{Case Study Protocols}
 
-\subsection{Secure Protocols Evaluated}
+For our evaluation, we implemented six MPC protocols from the literature in \system, and evaluated \system's scalability and ability to find implementation bugs in these protocols.
+The six protocols are listed in Table~\ref{tbl:protocols}.
+One is a simple addition protocol, three are basic building blocks often used to build more complex protocols, and two are protocols designed to securely evaluate circuits.
+All six protocols are designed for two parties, with the addition of a third ``dealer'' party to produce correlated randomness in several of the protocols.
 
@@
   \caption{Secure protocols used in our evaluation.}
+  \label{tbl:protocols}
 \end{table}
+
+\paragraph{$n$-bit Addition.}
+The simplest protocol in our evaluation is a basic summation protocol, which adds up the inputs from the two participating parties.
+To accomplish this task, the protocol essentially simulates a hardware adder circuit---for each bit in the binary representation of the two input numbers, the protocol performs bitwise operations to compute the sum and carry bit.
+To ensure security, the input values are \emph{secret shared} using additive secret sharing; to secret share a bit, a party selects a random bit to represent one share, and computes the other share by XORing the random bit with the secret bit.
+In \system, $P_1$ can secret share the bit $x$ with $P_2$ as follows:
+%
+\begin{verbatim}
+  s1 = FLIP @P1
+  s2 = x + s1
+  SEND s1 TO P2
+\end{verbatim}
+%
+$P_1$ can safely send one share to $P_2$, as long as $P_1$ keeps the other share for themselves---neither share alone reveals anything about the secret bit.
+To add (i.e.
+XOR) two secret-shared bits, each party can simply add its shares for the two bits to obtain one share of the result.
+To multiply (i.e.
+AND) two secret-shared bits, the parties must use oblivious transfer (described next).
+
+Our $n$-bit addition protocol generates a \system source file for a given bitwidth.
+In our evaluation, we test several different bitwidths to examine the effect of the size of each party's view on the scalability and bug-finding capability of \system.
+
+\paragraph{Oblivious Transfer.}
+Oblivious transfer~\cite{todo} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
+In 1-out-of-2 OT, $S$ provides 2 secret inputs $x_0$ and $x_1$ and $R$ provides a secret \emph{selection bit} $b$.
+If $b=0$, $R$ receives $x_0$; if $b=1$, $R$ receives $x_1$.
+$R$ does not learn anything about the not-selected secret, and $S$ does not learn which secret $R$ selected.
+This idea can be extended to 1-out-of-$n$ OT by extending the selection bit $b$ to a $\log_2(n)$-bit string.
+%
+Oblivious transfer is a building block for many other protocols. For example, the GMW protocol for circuit evaluation (described later) evaluates AND gates using 1-out-of-4 OT
+
+\paragraph{Oblivious Transfer Extension.}
+Protocols like GMW (described later) require the use of OT for each AND gate---for large circuits, this means thousands or millions of OTs.
+\emph{OT extension}~\cite{todo} is an approach for expanding $k$ ``base'' OTs into $m \gg k$ OTs.
+OT extension can significantly improve the performance of protocols that use many OTs.
+
+\paragraph{Beaver Triple Generation.}
+A \emph{Beaver triple}~\cite{todo} (or \emph{multiplication triple}) is a set of secret shares of numbers $a$, $b$, and $c$ such that $a \cdot b = c$.
+To use a Beaver triple, $P_1$ must have one share of each of $a$, $b$, and $c$, and $P_2$ must have the other share.
+Since the values are secret shared, they appear random to each party---but the correlation between the three numbers can be used to perform efficient multiplication between secret-shared numbers (described later).
+One Beaver triple must be consumed for each multiplication, but they can be generated in a separate offline phase.
+Beaver triples can be generated by a trusted ``dealer'' party, or by using oblivious transfer.
+We implement both versions in \system.
+
+\paragraph{GMW (Goldreich-Micali-Widgerson.}
+
+\paragraph{MPC with Beaver Triples.}
+
+\paragraph{Circuits.}
+To evaluate the circuit-based protocols, we used five binary circuits that are publicly available in the ``Bristol-fashion'' format~\cite{todo}.
+We chose commonly-used circuits with a range of sizes to test \system's scalability.
+Table~\ref{tbl:circuits} lists the circuits used in our evaluation, along with the number of gates in each circuit.
 
@@
   \caption{Circuits used in our evaluation.}
+  \label{tbl:circuits}
 \end{table}
+
+\section{Evaluation}
 

### 2023-10-01 python/circuit_experiment.py
let the circuit experiment skip stuff it's already done.

@@ -1,11 +1,93 @@
+from argparse import (ArgumentParser, FileType)
+from collections import Counter
+from csv import DictReader
+from dataclasses import dataclass
+from datetime import datetime
+from itertools import chain
+import math
 import os
+from pprint import PrettyPrinter
 import subprocess
 import sys
-import math
 
-from datetime import datetime
 
-now = datetime.now()
+launch_time = datetime.now()
+
+
+@dataclass(frozen=True)
+class SecurityConfiguration:
+    bias_sharing : int = 0
+    bias_and : int = 0
+    accidental_secret : float = 0
+    accidental_gate : float = 0
+    def __str__(self):
+        return f"{self.bias_sharing},{self.bias_and},{self.accidental_secret},{self.accidental_gate}"
+
+
+argp = ArgumentParser(description="Build various circuits and test them at various powers")
+# We don't use `FileType` arguments because we don't want these handles open all the time.
+argp.add_argument("--iters", "-i", type=int, action="extend", nargs="+", help="Values of \"iterations\" at which to test the circuits.")
+argp.add_argument("--trains", "-t", type=int, action="extend", nargs="+", help="Values of \"training N\" at which to test the circuits.")
+argp.add_argument("--testing-ratio", type=int, default=2,
+                  help="Circuits will be tested with \"testing N\" set to t/testing-ratio for each t in --trains.")
+argp.add_argument("--circuits", "-c", type=str, action="extend", nargs="+", help="Circuit files to parse into CHO and then test."
+                                                                                 +" A \"circuits/\" prefix will be added!")
+argp.add_argument("--protocols", "-p", type=str, action="extend", nargs="+", help="The protocols with which to implement each circuit.")
+argp.add_argument('--output', "-o", type=str, default="results/circuit_results_{t}.csv",
+                  help="Path to the CSV file to write data to. Will be overwritten if it already exists."
+                       +" The pattern \"\{t\}\" will be replaced with the current time.")
+argp.add_argument("--repititions", "-R", type=int, default=3, help="How many times to repeat each test on each circuit.")
+argp.add_argument("--cho-temp", type=str, default='tmp.cho', help="The temp-file to which to save the generated CHO protocol.")
+argp.add_argument("--pre-existing", "-P", type=str, action="extend", nargs="+", default=[],
[…]

### 2023-10-01 paper/main.tex
update definition

--- 2023-09-15 paper/main.tex
@@
 generated.
-The \termOfArt{ideal world view} of a party is what it sees in the ideal world: just its inputs and outputs.
-The \termOfArt{real world view} of a party is everything it sees in the real world while executing its part of Π:
+The \termOfArt{ideal world view}, $\mathsf{Ideal}_P(F,\vec{x})$, of a party P is what it sees when running F in the ideal world:
+just its inputs $\vec{x}_P$ and its outputs.
+The \termOfArt{real world view}, $\mathsf{Real}_P(Π, \vec{x})$, of a party P is everything it sees while executing its part of Π in the real world:
 its inputs, outputs, all messages it receives, and all the bits of random tape it uses.
+Both these concepts extend to naturally to lists of parties.
 
@@
     is secure against a passive adversary controlling C$\subset$P iff
-there exists a poly-time-computable function SIM s.t.
-....
+    there exists a poly-time-computable function $\mathtt{SIM}$ s.t.
+    $$\begin{aligned}
+        \mathsf{Real}_C(Π, \vec{x}) &\cong \mathtt{SIM}(\mathsf{Ideal}_C(F,\vec{x}))
+    \end{aligned}$$
+    where $\cong$ denotes distributional equivelence or indistinguishabilty.
 \end{definition}

### 2023-10-13 python/bristol_adder_generator.py
don't output the carry bit in adder circuits

@@ -69,9 +69,9 @@ for outv in outs:
     emit(f'1 1 {outv} {w} EQW')
     output_wires.append(w)
 
-w = next_wire()
-output_wires.append(w)
-emit(f'1 1 {carry} {w} EQW')
+# w = next_wire()
+# output_wires.append(w)
+# emit(f'1 1 {carry} {w} EQW')
 
 
 num_gates = f.getvalue().count('\n')

### 2023-10-15 python/gmw_circuit_generator.py
don't use accidental_gate for XOR gates.

@@ -26,8 +26,6 @@ def gen_xor(a, b):
     out = gensym('g')
     emit(f'{out}_1 = {a}_1 + {b}_1')
     emit(f'{out}_2 = {a}_2 + {b}_2')
-    if random.random() < config['accidental_gate']:
-        emit(f'SEND {out}_2 TO P1 -- accidental send to corrupt')
     return out
 
 def gen_inv(a):

### 2023-10-15 paper/main.tex
WIP on related-work section.

@@
 
-In practice, security is usually quantified over all possible corruptions C
-where |C|$\leq$|P|/x for some x, but this is not fundamental and we only consider a single C at a time.
+In practice, security is usually quantified over all possible corruptions $C$
+where $\lvert C \rvert \leq \frac{\lvert P \rvert}{x}$ for some $x$, but this is not fundamental and we only consider a single $C$ at a time.
 
@@
 That said, some simplification is possible for our purposes:
-Since the corrupted parties are assumed to follow Π faithfully, we are fine to limit ourselves to
+Since the (passive) corrupted parties are assumed to follow Π faithfully, we are fine to limit ourselves to
 choreographic protocols in which all messages sent are expected and all messages expected are received.
@@
 any protocol which is not \termOfArt{passive secure} is automatically not \termOfArt{active secure}.
+
+Certain edge cases are helpful to consider, as they really do map out the edges of the problem of testing from MPC security.
+In the first case, Π itself may be vacuous.
+A protocol in which no information (random bits don't count) is communicated is automatically secure,
+although of course it can only correctly implement trivial functionalities.
+On the other hand, if the adversary can efficiently deduce a unique possible honest input upon observing the corrupt output,
+then the security is vacuous; any correct implementation of that functionality is secure regardless of what's communicated.
 

### 2023-10-15 python/gmw_circuit_generator.py
Don't use accidental_gate for INV gates either.

@@ -32,8 +32,6 @@ def gen_inv(a):
     out = gensym('g')
     emit(f'{out}_1 = {a}_1 + 1')
     emit(f'{out}_2 = {a}_2')
-    if random.random() < config['accidental_gate']:
-        emit(f'SEND {out}_2 TO P1 -- accidental send to corrupt')
     return out
 
 def gen_and(a, b):

### 2023-10-15 python/beaver_circuit_generator.py
Accidential gate shouldn't result in stocastic circuit creation, and should apply to beaver the same amount as to gmw.

@@ -11,8 +11,8 @@ argp.add_argument("--bias_and", action="store", type=int, default=0,
                   help="Bias randomness used for AND gates")
 argp.add_argument("--accidental_secret", action="store", type=float, default=0.0,
                   help="Rate of accidentally sending secret inputs to corrupt party")
-argp.add_argument("--accidental_gate", action="store", type=float, default=0.0,
-                  help="Rate of accidentally sending shares of gate outputs to corrupt party")
+argp.add_argument("--accidental_gate", action="store", type=int, default=0,
+                  help="Rate of accidentally sending shares of and-gate outputs to corrupt party")
 args = argp.parse_args()
 
 config = {
@@ -38,19 +38,14 @@ def gen_and(a, b):
     out = gensym('g')
 
     emit(f'DO and_beaver(P1({a}_1, {b}_1), P2({a}_2, {b}_2)) GET({out}_1=out1, {out}_2=out2)')
-    if random.random() < config['accidental_gate']:
-        emit(f'SEND {out}_2 TO P1 -- accidental send to corrupt')
     return out
 
 generators = {'AND': gen_and, 'XOR': gen_xor, 'INV': gen_inv}
 
 share_randomness_var, share_randomness_defs = gen_randomness(config['bias_sharing'], 'P1')
-a1, a1_defs = gen_randomness(config['bias_and'], 'P3')
-a2, a2_defs = gen_randomness(config['bias_and'], 'P3')
-b1, b1_defs = gen_randomness(config['bias_and'], 'P3')
-b2, b2_defs = gen_randomness(config['bias_and'], 'P3')
+# we only bias one of the dealer flips, to try to keep the scale of this break on par with the other implementation.
 c1, c1_defs = gen_randomness(config['bias_and'], 'P3')
-c2, c2_defs = gen_randomness(config['bias_and'], 'P3')
+and_leakage_var, and_leakage_defs = gen_randomness(config['accidental_gate'], 'P2', baseline=0)
 
 header = f"""
 MACRO secret_share(P1(x), P2()) AS
@@ -62,14 +57,10 @@ ENDMACRO
 
[…]

### 2023-10-15 python/circuit_generator.py

@@ -1,11 +1,12 @@
 import random
 import sys
 
-def gen_randomness(bias_level, party):
+def gen_randomness(bias_level, party, baseline=1):
     output = ""
-    for i in range(bias_level+1):
+    for i in range(bias_level + baseline):
         output += f'f{i} = FLIP @{party}\n  '
-    exp = ' ^ '.join([f'f{i}' for i in range(bias_level+1)])
+    exp = ' ^ '.join([f'f{i}' for i in range(bias_level + baseline)]) or ' 1 '
+    # if baseline=bias_level=0, then you'll always use the value 1 instead of anding together the (nonexistant) flips.
     return exp, output
 
 gn = 0

### 2023-10-15 python/gmw_circuit_generator.py

@@ -11,8 +11,8 @@ argp.add_argument("--bias_and", action="store", type=int, default=0,
                   help="Bias randomness used for AND gates")
 argp.add_argument("--accidental_secret", action="store", type=float, default=0.0,
                   help="Rate of accidentally sending secret inputs to corrupt party")
-argp.add_argument("--accidental_gate", action="store", type=float, default=0.0,
-                  help="Rate of accidentally sending shares of gate outputs to corrupt party")
+argp.add_argument("--accidental_gate", action="store", type=int, default=0,
+                  help="Rate of accidentally sending shares of and-gate outputs to corrupt party")
 args = argp.parse_args()
 
 config = {
@@ -37,14 +37,13 @@ def gen_inv(a):
 def gen_and(a, b):
     out = gensym('g')
     emit(f'DO and_gmw(P1({a}_1, {b}_1), P2({a}_2, {b}_2)) GET({out}_1=out1, {out}_2=out2)')
-    if random.random() < config['accidental_gate']:
-        emit(f'SEND {out}_2 TO P1 -- accidental send to corrupt')
     return out
 
 generators = {'AND': gen_and, 'XOR': gen_xor, 'INV': gen_inv}
 
 share_randomness_var, share_randomness_defs = gen_randomness(config['bias_sharing'], 'P1')
 and_randomness_var, and_randomness_defs = gen_randomness(config['bias_and'], 'P2')
+and_leakage_var, and_leakage_defs = gen_randomness(config['accidental_gate'], 'P2', baseline=0)
 
 header = f"""
 MACRO secret_share(P1(x), P2()) AS
@@ -57,6 +56,9 @@ ENDMACRO
 MACRO and_gmw(P1(x2, y2), P2(x1, y1)) AS
   {and_randomness_defs}
   out2 = {and_randomness_var}
+  {and_leakage_defs}
+  leakage = out2 ^(~({and_leakage_var}))
+  SEND leakage TO P1
   g1_s2_00 = out2 + ((x1 + 0) ^ (y1 + 0))
   g1_s2_01 = out2 + ((x1 + 0) ^ (y1 + 1))
   g1_s2_10 = out2 + ((x1 + 1) ^ (y1 + 0))

### 2023-10-16 python/circuit_generator.py
accidental_secret also results in deterministic .cho generation

@@ -58,8 +58,10 @@ def gen_circuit(config, generators, header, circuit_file, output_file):
     emit()
     for yi in ys:
         emit(f'{yi} = SECRET @P2')
-        if random.random() < config['accidental_secret']:
-            emit(f'SEND {yi} TO P1 -- accidental send secret to corrupt')
+        value, setup = gen_randomness(config['accidental_secret'], 'P2', baseline=0)
+        emit(setup)
+        emit(f'leak_{yi} = {yi} ^(~({value}))')
+        emit(f'SEND leak_{yi} TO P1 -- accidental send secret to corrupt')
 
     emit()
     emit('-- Set up shares')

### 2023-10-16 python/beaver_circuit_generator.py
change all-perfect failsafe to 0.5; finish up argument transition to ints.

@@ -9,7 +9,7 @@ argp.add_argument("--bias_sharing", action="store", type=int, default=0,
                   help="Bias randomness used in secret sharing")
 argp.add_argument("--bias_and", action="store", type=int, default=0,
                   help="Bias randomness used for AND gates")
-argp.add_argument("--accidental_secret", action="store", type=float, default=0.0,
+argp.add_argument("--accidental_secret", action="store", type=int, default=0,
                   help="Rate of accidentally sending secret inputs to corrupt party")
 argp.add_argument("--accidental_gate", action="store", type=int, default=0,
                   help="Rate of accidentally sending shares of and-gate outputs to corrupt party")

### 2023-10-16 python/circuit_experiment.py

@@ -18,8 +18,8 @@ launch_time = datetime.now()
 class SecurityConfiguration:
     bias_sharing : int = 0
     bias_and : int = 0
-    accidental_secret : float = 0
-    accidental_gate : float = 0
+    accidental_secret : int = 0
+    accidental_gate : int = 0
     def __str__(self):
         return f"{self.bias_sharing},{self.bias_and},{self.accidental_secret},{self.accidental_gate}"
 
@@ -98,8 +98,8 @@ configs = [
     SecurityConfiguration(), # secure
     *( SecurityConfiguration(bias_sharing      = i  ) for i in range(1, 4) ),  # for consistency i should make these args too, but it's a chore.
     *( SecurityConfiguration(bias_and          = i  ) for i in range(1, 4) ),
-    *( SecurityConfiguration(accidental_secret = 1/i) for i in range(2, 5) ),
-    *( SecurityConfiguration(accidental_gate   = 1/i) for i in range(2, 5) )
+    *( SecurityConfiguration(accidental_secret = i  ) for i in range(1, 4) ),
+    *( SecurityConfiguration(accidental_gate   = i  ) for i in range(1, 4) )
 ]
 log(configs)
 
@@ -114,8 +114,8 @@ def scrape_prior(csv_name):
                    int(row["test_size"]),
                    SecurityConfiguration(bias_sharing=int(row["bias_sharing"]),
                                          bias_and=int(row["bias_and"]),
-                                         accidental_secret=float(row["accidental_secret"]),
-                                         accidental_gate=float(row["accidental_gate"])))
+                                         accidental_secret=int(row["accidental_secret"]),
+                                         accidental_gate=int(row["accidental_gate"])))
 
 already_done = Counter(chain.from_iterable(scrape_prior(p) for p in args.pre_existing))
 log(already_done)

### 2023-10-16 python/d-tree-csv.py

@@ -108,7 +108,7 @@ def statistical_test(a, b) -> float:
         except ValueError as ve:
             message, *_ = ve.args
             if message == "zero_method 'wilcox' and 'pratt' do not work if x - y is zero for all elements.":
-                return 1  # If all the differences are zero, then we have zero evidence of insecurity.
+                return 0.5  # If all the differences are zero, then we have zero evidence of insecurity.
             else:
                 raise ve
 

### 2023-10-17 python/beaver_circuit_generator.py
fix errant text

@@ -88,7 +88,7 @@ MACRO and_beaver(P1(x2, y2), P2(x1, y1)) AS
   e = e1 + e2
 
   out1 = (d ^ e) + (d ^ b1) + (e ^ a1) + c1
-  out2 = (d ^ e) y or have a sp+ (d ^ b2) + (e ^ a2) + c2
+  out2 = (d ^ e) + (d ^ b2) + (e ^ a2) + c2
 
   {and_leakage_defs}
   leakage = out2 ^(~({and_leakage_var}))

### 2023-10-17 python/circuit_generator.py
modify generated circuits to not output anything

@@ -126,10 +126,10 @@ def gen_circuit(config, generators, header, circuit_file, output_file):
     output_names = [wire_names[k] for k in output_wires]
 
     emit()
-    emit('-- Reveal output')
-    rs = [f'r{i}' for i in range(len(output_names))]
-    for o, r in zip(output_names, rs):
-        emit(f'DO reveal(P1({o}_1), P2({o}_2)) GET({r}=y)')
-        emit(f'OUTPUT {r}')
+    # emit('-- Reveal output')
+    # rs = [f'r{i}' for i in range(len(output_names))]
+    # for o, r in zip(output_names, rs):
+    #     emit(f'DO reveal(P1({o}_1), P2({o}_2)) GET({r}=y)')
+    #     emit(f'OUTPUT {r}')
 
     global_f.close()

### 2023-10-26 python/bristol_lt_generator.py
hopefully fix lt generator

@@ -27,6 +27,15 @@ def gen_xor(a, b):
     emit(f'2 1 {a} {b} {out} XOR')
     return out
 
+def gen_or(a, b):
+    out1 = next_wire()
+    out2 = next_wire()
+    out = next_wire()
+    emit(f'2 1 {a} {b} {out1} AND')
+    emit(f'2 1 {a} {b} {out2} xor')
+    emit(f'2 1 {out1} {out2} {out} XOR')
+    return out
+
 def gen_and(a, b):
     out = next_wire()
     emit(f'2 1 {a} {b} {out} AND')
@@ -34,18 +43,13 @@ def gen_and(a, b):
 
 def gen_less_than(a_bits, b_bits):
     a_greater = None
-    b_greater = None
 
     for a, b in zip(a_bits, b_bits):
         if a_greater is None:
-            assert b_greater is None
             new_a_greater = gen_and(a, gen_not(b))
-            new_b_greater = gen_and(b, gen_not(a))
         else:
-            new_a_greater = gen_xor(a_greater, gen_and(a, gen_and(gen_not(b), gen_not(b_greater))))
-            new_b_greater = gen_xor(b_greater, gen_and(b, gen_and(gen_not(a), gen_not(a_greater))))
+            new_a_greater = gen_or(a_greater, gen_and(a, gen_not(b)))
         a_greater = new_a_greater
-        b_greater = new_b_greater
     return a_greater
 
 

### 2023-10-26 python/cho_builder/beaver.py
fix beaver triple-based implementation

@@ -46,7 +46,8 @@ MACRO and_beaver(P1(x2, y2), P2(x1, y1)) AS
   e = e1 + e2
 
   out1 = (d ^ e) + (d ^ b1) + (e ^ a1) + c1
-  out2 = (d ^ e) + (d ^ b2) + (e ^ a2) + c2
+  -- only one party does addition-w-constant.
+  out2 =           (d ^ b2) + (e ^ a2) + c2
 
   {and_leakage_defs}
   leakage = out2 ^(~({and_leakage_var}))

### 2023-10-26 python/bristol_lt_generator.py
fix lowercase xor

@@ -32,7 +32,7 @@ def gen_or(a, b):
     out2 = next_wire()
     out = next_wire()
     emit(f'2 1 {a} {b} {out1} AND')
-    emit(f'2 1 {a} {b} {out2} xor')
+    emit(f'2 1 {a} {b} {out2} XOR')
     emit(f'2 1 {out1} {out2} {out} XOR')
     return out
 

### 2023-10-26 python/bristol_lt_generator.py
one more try at lt circuits

@@ -43,13 +43,18 @@ def gen_and(a, b):
 
 def gen_less_than(a_bits, b_bits):
     a_greater = None
+    b_greater = None
 
     for a, b in zip(a_bits, b_bits):
         if a_greater is None:
+            assert b_greater == None
             new_a_greater = gen_and(a, gen_not(b))
+            new_b_greater = gen_and(b, gen_not(a))
         else:
-            new_a_greater = gen_or(a_greater, gen_and(a, gen_not(b)))
+            new_a_greater = gen_or(a_greater, gen_and(a, gen_and(gen_not(b), gen_not(b_greater))))
+            new_b_greater = gen_or(b_greater, gen_and(b, gen_and(gen_not(a), gen_not(a_greater))))
         a_greater = new_a_greater
+        b_greater = new_b_greater
     return a_greater
 
 

### 2023-10-26 python/cho_builder/beaver.py
change the beaver useage of bias_and again

@@ -15,12 +15,13 @@ generators = GateGenerators(and_gate=gen_and, xor_gate=gen_xor, inv_gate=gen_inv
 and_macro = """
 MACRO and_beaver(P1(x2, y2), P2(x1, y1)) AS
   -- generate beaver triple
-  a1 = FLIP @P3
+  {a1_defs}
+  a1 = {a1}
   a2 = FLIP @P3
   b1 = FLIP @P3
-  b2 = FLIP @P3
-  {c1_defs}
-  c1 = {c1}
+  {b2_defs}
+  b2 = {b2}
+  c1 = FLIP @P3
 
   c2 = ((a1 + a2) ^ (b1 + b2)) + c1
 
@@ -65,14 +66,17 @@ Broken acording to: {config}
 def main(input_file, output_file, config):
     share_randomness_var, share_randomness_defs = gen_randomness(config.bias_sharing, 'P1')
     # we only bias one of the dealer flips, to try to keep the scale of this break on par with the other implementation.
-    c1, c1_defs = gen_randomness(config.bias_and, 'P3')
+    a1, a1_defs = gen_randomness(config.bias_and, 'P3', globalizer='a')
+    b2, b2_defs = gen_randomness(config.bias_and, 'P3', globalizer='b')
     and_leakage_var, and_leakage_defs = gen_randomness(config.accidental_gate, 'P2', baseline=0)
 
     header = "\n".join((header_comment.format(input_file=input_file, config=config),
                         share_macro.format(share_randomness_var=share_randomness_var,
                                            share_randomness_defs=share_randomness_defs),
-                        and_macro.format(c1=c1,
-                                         c1_defs=c1_defs,
+                        and_macro.format(a1=a1,
+                                         a1_defs=a1_defs,
+                                         b2=b2,
+                                         b2_defs=b2_defs,
                                          and_leakage_var=and_leakage_var,
                                          and_leakage_defs=and_leakage_defs),
                         reveal_macro.format()))

### 2023-10-29 paper/main.tex
updates to evaluation

@@
 \maketitle
+\thispagestyle{plain}
+\pagestyle{plain}
 
@@
     Oblivious Transfer & Building block & TODO \\
-    OT Extension & Building block & TODO \\
+%    OT Extension & Building block & TODO \\
     Beaver Triple Generation & Building block & TODO \\
@@
 %
-Oblivious transfer is a building block for many other protocols. For example, the GMW protocol for circuit evaluation (described later) evaluates AND gates using 1-out-of-4 OT
+Oblivious transfer is a building block for many other protocols.
 
-\paragraph{Oblivious Transfer Extension.}
-Protocols like GMW (described later) require the use of OT for each AND gate---for large circuits, this means thousands or millions of OTs.
+% \paragraph{Oblivious Transfer Extension.}
+% Protocols like GMW (described later) require the use of OT for each AND gate---for large circuits, this means thousands or millions of OTs.
 \emph{OT extension}~\cite{todo} is an approach for expanding $k$ ``base'' OTs into $m \gg k$ OTs.
@@
 
-\paragraph{GMW (Goldreich-Micali-Widgerson.}
+\paragraph{GMW (Goldreich-Micali-Widgerson).}
+The GMW protocol is an MPC protocol for evaluating circuits.
+Each party holds additive secret shares of each wire's value, and evaluates addition gates by adding the shares they hold.
+The parties evaluate multiplication gates by using 1-out-of-4 oblivious transfer: the first party generates a random bit for its share of the output, and acts as the sender in OT; the second party acts as the receiver in OT and receives its share of the output.
+We implemented a compiler to generate \system programs that evaluate  Bristol fashion circuits using GMW, and evaluated it on circuits implementing integer addition and integer inequalities.
 
 \paragraph{MPC with Beaver Triples.}
+Another MPC protocol for evaluating circuits also uses additive shares to hold wire values, but uses Beaver triples to evaluate multiplication gates (instead of OT).
+Given shares of a Beaver triple, the parties perform two multiplication-by-constant and broadcast operations, and derive one share each of the product.
+Our implementation of this approach employees a third-party ``dealer'' who generates the triples; a pre-processing (offline) phase can also be used to generate the triples.
+We evaluate \system on this protocol using the same circuits as in our GMW experiments.
 
-\paragraph{Circuits.}
-To evaluate the circuit-based protocols, we used five binary circuits that are publicly available in the ``Bristol-fashion'' format~\cite{todo}.
+% \paragraph{Circuits.}
+% To evaluate the circuit-based protocols, we used five binary circuits that are publicly available in the ``Bristol-fashion'' format~\cite{todo}.
 We chose commonly-used circuits with a range of sizes to test \system's scalability.
@@
 
-\begin{table}
-\centering
-  \begin{tabular}{|l |l|l|}
-    \hline
-    \textbf{Circuit} & \textbf{AND Gates} & \textbf{Total Gates} \\
-    \hline
-    64-bit adder & 63 & 376 \\
-    64-bit multiplier & 4,033 & 13,675 \\
-    AES-128 & 6,400 & 36,663 \\
-    SHA-256 & 22,573 & 135,073 \\
-    Floating-point division & 82,269 & 184,007 \\
-    \hline
-  \end{tabular}
-  \caption{Circuits used in our evaluation.}
-  \label{tbl:circuits}
-\end{table}
+% \begin{table}
+% \centering
+%   \begin{tabular}{|l |l|l|}
+%     \hline
+%     \textbf{Circuit} & \textbf{AND Gates} & \textbf{Total Gates} \\
+%     \hline
+%     64-bit adder & 63 & 376 \\
+%     64-bit multiplier & 4,033 & 13,675 \\
+%     AES-128 & 6,400 & 36,663 \\
+%     SHA-256 & 22,573 & 135,073 \\
+%     Floating-point division & 82,269 & 184,007 \\
+%     \hline
+%   \end{tabular}
+%   \caption{Circuits used in our evaluation.}
+%   \label{tbl:circuits}
+% \end{table}
 
@@
 
-\subsection{Experiment 1: Scalability for Secure Protocols}
+\begin{itemize}
+\item \textbf{RQ1:} Is \system capable of detecting insecurity in complex, multi-round protocols?
+\item \textbf{RQ2:} Is \system capable of detecting insecurity in real-world protocols?
+\end{itemize}
 
-\subsection{Experiment 2: Bug-Finding for Insecure Protocol Mutations}
+\subsection{Experiment 1: Real-world Secure Protocols}
 
-\subsection{Experiment 3: Scalability and Bug-Finding for Random Programs}
+\subsection{Experiment 2: Random Protocols}
 

### 2023-10-29 paper/main.tex
skeleton of the evaluation section

@@
 \section{Case Study Protocols}
+\label{sec:case-study-protocols}
 
@@
 
+Our empirical evaluation demonstrates \system's ability to detect bugs in secure protocols. Our experiments are designed to answer two research questions:
+%
 \begin{itemize}
@@
 \end{itemize}
+%
+To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \system's ability to show they are insecure.
+The details appear in Section~\ref{sec:experiment-1:-random}.
+To answer RQ2, we evaluate \system on the case study protocols described in Section~\ref{sec:case-study-protocols}.
+The details appear in Section~\ref{sec:experiment-2:-real}.
+In both cases, \system is able to detect insecurity in the target protocols in under a minute.
 
-\subsection{Experiment 1: Real-world Secure Protocols}
+\subsection{Experiment 1: Random Protocols}
+\label{sec:experiment-1:-random}
 
-\subsection{Experiment 2: Random Protocols}
+\paragraph{Experiment setup.}
+
+\paragraph{Results.}
+
+\subsection{Experiment 2: Real-world Protocols}
+\label{sec:experiment-2:-real}
+
+\paragraph{Experiment setup.}
+
+\paragraph{Results.}
 

### 2023-10-29 paper/main.tex
abstract

--- 2023-10-15 paper/main.tex
@@
 \newunicodechar{Π}{\ifmmode\Pi\else\textPi\fi}
+
+% use quotes as smart-quotes. https://tex.stackexchange.com/a/52354
+\usepackage [english]{babel}
+\usepackage [autostyle, english = american]{csquotes}
+\MakeOuterQuote{"}
 
@@
 \begin{abstract}
-  Abstract goes here.
+  Formal methods for guaranteeing that a protocol satisfies a cryptographic security definition have advanced substantially,
+  but such methods are still labor intensive and the need remains for an automated tool that can positively identify an insecure protocol.
+  We show that the philosophy of property-based-testing, \textit{"run it a bunch of times and see if it breaks"},
+  is at least somewhat feasible in a cryptographic setting.
+  We specifically target Secure Multi-Party Computation, because formal methods targeting this security definition for bit-model implementations
+  are particularly difficult.
+  Using results from the literature for Probabilistic Programming Languages and statistical inference, we devise a test that can
+  detect various flaws in a bit-level implementation of an MPC protocol.
+  The test is grey-box; it requires only transcripts of randomness consumed by the protocol and of the inputs, outputs, and messages.
+  It successfully detects several different mistakes and biases introduced into two different implementations of the classic GMW protocol.
+  Applied to thousands of randomly generated protocols, it identifies nearly all of them as insecure.
+  We also include an analysis of the parameters of the test, and discussion of what makes detection of MPC (in)security difficult.
 \end{abstract}
@@
 
-In the work we particularly focus on Secure Multi-Party Computation (MPC),
+In this work we particularly focus on Secure Multi-Party Computation (MPC),
 a cryptographic primitive with wide-ranging applications but frustrating performance limitations.

### 2023-10-29 paper/main.tex
background is fully drafted

@@
 a cryptographic primitive with wide-ranging applications but frustrating performance limitations.
-Substantial work has going into improving the performance of MPC systems [citation].
+Substantial work has going into improving the performance of MPC systems [TODO: citation].
 Additionally, in some cases performance can be improved by using a bespoke protocol instead of a general-purpose MPC system [do we have any evidence of this?].
@@
 Implementations that work on arbitrary functions (representable as circuit computations in some field)
-have existed since [...],
+have existed since 1982 [TODO: citation], % https://ieeexplore.ieee.org/document/4568207
 but their use by the public is limited by their performance.
@@
 from all other parties, compute F, and then send the parties their respective outputs.
-A \termOfArt{real world} protocol Π is said to correctly implement F if, for all possible inputs,
+A \termOfArt{real world} protocol Π is said to \textit{correctly} implement F if, for all possible inputs,
 Π results in all parties getting the same outputs as they would have in the ideal world.
-Defining the security of Π requires a clear threat model, and requires defining the participants' "views".
+Defining the \textit{security} of Π requires a clear threat model, and requires defining the participants' "views".
 
@@
 This adversary cannot cause the parties to deviate from the protocol Π in any way;
-we are interested in whether it can learn anything beyond the outputs of the "corrupted" machines
+we are interested in whether it can learn anything (beyond the outputs of the "corrupted" machines)
 by observing the execution of Π.
@@
 In practice, security is usually quantified over all possible corruptions $C$
-where $\lvert C \rvert \leq \frac{\lvert P \rvert}{x}$ for some $x$, but this is not fundamental and we only consider a single $C$ at a time.
+where $\lvert C \rvert \leq \frac{\lvert P \rvert}{r}$ for some $r$, but this is not fundamental and we only consider a single $C$ at a time.
 
@@
 Since the (passive) corrupted parties are assumed to follow Π faithfully, we are fine to limit ourselves to
-choreographic protocols in which all messages sent are expected and all messages expected are received.
+\termOfArt{choreographic} protocols in which all messages sent are expected and all messages expected are received.
 In this context, and because the only source of non-determinism is the parties' random tapes
@@
 Then we can collapse all the corrupted parties C into a single party C,
-and similarly all the other parties into a single party H (for "honest").
+and similarly all the other parties into a single party H (for "honest")
+(we don't always do this).
 Finally, in this work we assume that protocols implement their functionalities \textit{correctly},
@@
 In the first case, Π itself may be vacuous.
-A protocol in which no information (random bits don't count) is communicated is automatically secure,
+A \termOfArt{vacuous protocol}, one in which no information (random bits don't count), is communicated is automatically secure,
 although of course it can only correctly implement trivial functionalities.
-On the other hand, if the adversary can efficiently deduce a unique possible honest input upon observing the corrupt output,
-then the security is vacuous; any correct implementation of that functionality is secure regardless of what's communicated.
+On the other hand, if the ideal-world adversary can efficiently deduce a unique possible honest input upon observing the corrupt output,
+then the protocol is \termOfArt{vacuously secure}: \textit{any} correct implementation of that functionality is secure regardless of what's communicated.
+These extremes mark the ends of a spectrum; most protocols are somewhere in the middle,
+where learning the output focuses the corrupt party's distribution-estimate of the honest secrets but still leaves some uncertainty.
+On the other hand, this spectrum doesn't capture the assumptions about the adversary's computational limits;
+if inverting the functionality F is NP-hard, then the protocol is \textit{not} vacuously secure, even if F is a bijection.
 
-[Other aspects we may wish to cover:
- vacuous protocols,
- vacuous security,
- infertabilty of functionality,
- comparison to control flow properties,
- review of other work,
- review of stuff we tried that didn't work...
- ]
+While MPC can be expressed as a kind of non-interference property, this is only somewhat useful.
+The non-interference can only be meant within the space of possible program traces \textit{posterior to the output}.
+Similarly, trying to avoid such subtleties in the non-interference property by providing a correctly constrained \texttt{reveal} operator
+isn't promising because constraints on \texttt{reveal} would need to recapitulate the original question of security
+and examine the entire rest of the protocol as context.
+
+\subsection{Inference in PPLs}
+
+Efficient inversion of programs and fragments of programs turns out to be so important for determining if a protocol satisfies MPC security.
+Since MPC protocols are probabilistic, what we really need is efficient posterior inference of program inputs in a
+\termOfArt{probabilistic programming language} (PPL).
+
+A PPL is a language that provides stochastic non-deterministic behavior as a language primitive
+and provides meta-theory or auxiliary tooling to reason about the resulting distributions of program behavior.
+In particular, PPLs commonly have semantics defined in terms of transformations on distributions,
+and are designed to ensure that the resulting distributions are amenable to specified techniques for efficient
+inference and sampling. Since even representing a fully-general multivariate distribution takes exponential space with respect to the number of variables,
+PPLs typically restrict the programs and/or distributions they can handle.
+
+For application to the subject of MPC protocols, a PPL would need to handle the two different kinds of uncertainty facing a hypothetical adversary:
+1) Random values generated (by any party) at runtime have known distributions
+(\eg the python expression \texttt{secrets.SystemRandom().random()}
+will sample at runtime from the best floating-point approximation the machine can provide
+of the uniform distribution on $\mathbb{R}_{(0,1]}$).
+2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
+possible secrets representing their prior belief so that the secret inputs can also be modeled as random variables.
+
+Hypothetically, supposing a protocol in a PPL providing a multivariate distribution over an appropriate subset of all runtime values,
+if it were possible to efficiently infer and sample from the real-world views
+(or other variables sufficient to deterministically compute a view)
+\textit{posterior to} the corrupt outputs (and corrupt inputs),
+then that would suffice as a simulator
+\textit{if and only if} the inferred distribution were independent of the adversary's choice of prior distribution for the honest secrets.
+So a direct test of MPC security (or a useful over-approximation) could be built from
+A) a language meta-theory guarantee that such posterior inference and sampling would be efficient,
+B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
+of the prior distribution of honest secrets[TODO: footnote], and
+C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs, presumably by handling them symbolically.
+Such a system is not necessarily impossible, but we have so far failed to build it.
+
+\subsection{Property-based testing}
+
+In industry the normal way to see if a program satisfies some desired property is to run it and see what it does.
+Sometimes this is done with a select handful of inputs or configurations,
+but non-deterministic programs are better evaluated with \termOfArt{property-based testing},
+in which the program is run with many arbitrary (but not necessarily random) configurations.
+Property-based testing is a popular paradigm for testing code in industry;
+the classic Haskell framework QuickCheck [citation] % https://www.cse.chalmers.se/~rjmh/QuickCheck/
+is still in use and has been ported to various other languages.
+
+At face value, property-based testing doesn't generally extend to hyper-properties,
+which can only be assessed by comparing multiple program traces.
+In practice sometimes one can express a hyper-property as a property of multiple of duplicate programs,
+but the sensitivity of the test quickly decays and the logistics of setting up the test quickly grow.
+Furthermore, when the property to be tested is over distributions of traces,
+the various used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
+
+\section{Related Work}
 
@@
   probabilistic separation logic]
-
-\subsection{Inference in PPLs}
-
-\subsection{Property-based testing}
-
-\section{Related Work}
-
 \subsection{Verification of MPC}

### 2023-10-30 paper/main.tex
comments and sketching of evaluation

@@
 
+TODO: maybe say we invented these; give examples. Readers will assume these are existing terms if we don't say otherwise.
+
 While MPC can be expressed as a kind of non-interference property, this is only somewhat useful.
@@
 and examine the entire rest of the protocol as context.
+
+TODO: maybe state it explicitly with the conditioning? Then we can say that no system can capture that property.
 
@@
 
+TODO: Is efficient inference really necessary? The conditional independence test might be sufficient? I think we don't want to imply that inference is sufficient, because lots of work has done that
+
 \subsection{Property-based testing}
@@
 the various used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
+
+TODO: maybe cite the Kifer work on property-based testing for DP (also a hyperproperty) and call out that our setting is (maybe) fundamentally high-dimensional so it's super tricky.
 
@@
 \item \textbf{RQ2:} Is \system capable of detecting insecurity in real-world protocols?
+\item \textbf{RQ3:} Is \system scalable to realistic protocols in reasonable time?
 \end{itemize}
@@
 \paragraph{Experiment setup.}
+In our first experiment, we generate random protocols and evaluate \system's ability to detect insecurity in these protocols.
+We randomly generate protocols according to the grammar of XXX.
+We bias the generated protocols in favor of generating and using randomness in each round of the protocol, in addition to the secret inputs.
+Random protocols of non-trivial length are extremely likely to be insecure.
+%
+We generate XXXX programs of size XXXX and evaluate \system's ability to detect insecurity.
 
 \paragraph{Results.}
+\emph{results description goes here}
 
@@
 \paragraph{Experiment setup.}
+In our second experiment, we evaluate \system's ability to detect insecurity in the real-world protocols described in Section~\ref{sec:case-study-protocols}.
+We run \system on the unmodified protocol in order to establish that the protocol is secure.
+Then, we mutate the protocol in several ways to introduce insecurity, and test \system's ability to detect it.
+Specifically, our mutations are:
+%
+\begin{itemize}
+\item \textbf{Accidental secret}: accidentally sends a secret held by the honest party  to the corrupt party.
+\item \textbf{Accidental gate}: accidentally sends an honest party's result from the evaluation of a multiplication gate to the corrupt party.
+\item \textbf{Biased sharing}: uses biased randomness to construct secret shares of honest party inputs.
+\item \textbf{Biased AND}: uses biased randomness as part of the evaluation of multiplication gates.
+\end{itemize}
+%
+The severity of each mutation can be tuned by adjusting the bias in the related randomness.
+In the case of ``accidental secret'' and ``accidental gate,'' this affects the probability of accidentally revealing a value; in the others, it directly adjust the bias in the randomness.
 
-\paragraph{Results.}
+\paragraph{Results.} \emph{results description goes here}
+
+\subsection{Discussion}
 

### 2023-10-30 paper/main.tex
prior work WIP :

--- 2023-10-29 paper/main.tex
@@
 
+Formal methods exist for verifying MPC protocols.
+\cite{https://arxiv.org/pdf/1806.07197.pdf} % Computer-aided proofs for multiparty computation with active security
+use EasyCrypt to show the security of GMW,
+but their technique is only applicable to a protocol with efficiently invertible functionality. 
+and
+\cite{Wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
+give fine-grain integration of 
 [ computer proofs of MPC paper; (double-check malicions case...)
@@
 \subsection{Verification of MPC}
-
-[Computer-aided proofs for multiparty computation with active security](https://arxiv.org/pdf/1806.07197.pdf)
 

### 2023-10-31 paper/main.tex
prior work WIP

--- 2023-10-30 paper/main.tex
@@
 
-[ computer proofs of MPC paper; (double-check malicions case...)
-  l-obliv
-  wysteria et al
-  probabilistic separation logic]
-\subsection{Verification of MPC}
+\subsection{Formal-methods for cryptography}
+Although a useful language-driven test for MPC security remains out of reach,
+a variety of useful work aproaches the subject from different directions.
 
-[Computer-aided proofs for multiparty computation with active security](https://arxiv.org/pdf/1806.07197.pdf)
+\cite{https://arxiv.org/pdf/1806.07197.pdf} % Computer-aided proofs for multiparty computation with active security
+use EasyCrypt to manually show the security of GMW.
+Their definitions generalize to protocols with efficiently invertible functionalities.
+% remove? Since existing general-purpose SMPC protocols can easily compute non-invertible functions, this is a serious limitation.
+They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
+If their system could be automated and paired with language constraints to ensure invertibility then it
+could apply to many, but certainly not all, use cases.
 
-\subsection{Languages built inside of MPC implementations}
+\cite{Barth et al} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
+give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
+Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.
+
+\cite{Wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
+give fine-grain integration of MPC as a language primitive, to help a larger program use it efficiently.
+[TODO: there are other/later works along these lines we should cite?]
+
+\cite{lambda-obliv} %
+gives a type system that ensures uniform and independent randomness of key values throughout a program.
+Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
+but this would of course prevent any non-vacuous functionality from being implemented.
 
@@
 
-dice
+We experimented with choreography languages that translate directly into Sum Product Networks (SPNs) \cite{TODO}.
+A SPN is directed-graph representation of a multi-variate distribution;
+they are compact, allow efficient inference, and can represent values symbolically as needed,
+so testing for conditional independence is practical.
+However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
+because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.
 
-\subsection{PPL proof assistants}
+The \cite{DICE} language analogously translates programs into binary decision diagrams (BDDs).
+BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
+but since the graph \textit{structure} of the BDD of a DICE program depends on the actual \textit{values} on which one is conditioning,
+a dice-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
 
-lilac
+\cite{lilac} % https://arxiv.org/abs/2304.01339
+is a proof-assistant system for reasoning about probability distributions and PPLs.
+It
 

### 2023-10-31 paper/main.tex
related work section is cluttered, but i think it mentions everyone...

@@
 
-TODO: maybe cite the Kifer work on property-based testing for DP (also a hyperproperty) and call out that our setting is (maybe) fundamentally high-dimensional so it's super tricky.
-
 \section{Related Work}
@@
 
+\cite{Gancher et al} % https://dl.acm.org/doi/pdf/10.1145/3571223
+give a powerful language system that can check if one protocol is a valid simulator of another,
+and show that it works on existing MPC protocols like GMW.
+The user of the system needs to write a second protocol
+representing the composition of the ideal-world functionality and the explicit function $\mathtt{SIM}$,
+and then build a high-level proof of the equivalence in the provided equational logic.
+
 \cite{Wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
@@
 but this would of course prevent any non-vacuous functionality from being implemented.
+
+\cite{OWL} % https://eprint.iacr.org/2023/473.pdf
+gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
+They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
+there's no way to verify an implementation of MPC.
+
+\cite{Fournet et al} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
+also use a type system to ensure safe usage of cryptographic primitives,
+but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
+Like OWL, their type system assumes a secure implementation of the encryption scheme.
 
@@
 is a proof-assistant system for reasoning about probability distributions and PPLs.
-It
+It more expressive than the original probabilistic separation logic,
+but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
 
-\subsection{Probabilistic tests of distributional equality}
+\subsection{Probabilistic testing}
 
-[decision trees paper]()
+\cite{Kifer} % https://arxiv.org/pdf/1805.10277.pdf
+and more recently \cite{Zhang et al} % https://dl.acm.org/doi/10.1145/3428233
+use large samples of program executions to test for violations of differential privacy,
+showing that such techniques can work on cryptographic distributional hyper-properties.
+Since differential privacy is only concerned with program inputs and outputs,
+the dimensionality of the problem is somewhat smaller than MPC.
+
+Our work builds directly on \cite{decision trees paper} % where did it go? all i can find is https://arxiv.org/pdf/2007.12325.pdf which definitely isn't what we worked off of...
+who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
+The following section explains this idea, and how we use it, in more detail.
 

### 2023-11-01 paper/main.tex
WIP: describing the tool

@@
 \usepackage{xspace}
+\usepackage{centernot}
 
@@
 \newunicodechar{Π}{\ifmmode\Pi\else\textPi\fi}
+\newunicodechar{⫫}{\mathrel{\perp\!\!\!\perp}}
 
@@
     there exists a poly-time-computable function $\mathtt{SIM}$ s.t.
+    for all combinations of inputs $\vec{x}$
     $$\begin{aligned}
@@
 who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
-The following section explains this idea, and how we use it, in more detail.
+Specifically, if decision trees (D-Trees trained on random variables $A$ and $C$ do better at predicting the value of a random variable $B$,
+then $A$ and $B$ are not independent conditioned on $C$.
+The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
+In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
+but in our adaptation we use fresh program traces for everything.
 
-\section{The tool in theory}
+\section{Statistical testing for MPC insecurity}
 
-... Position the equivalence in the definition of MPC security...
+\subsection{Test algorithm}
+
+If a simulator $\mathtt{SIM}$ exists for a protocol Π implementing functionality F,
+and we suppose fixed corrupt inputs $\vec{x}_C$,
+then for any observed $\vec{y}_C$ the values yielded by $\mathtt{SIM}(\vec{x}_C, \vec{y}_C)$
+will be drawn from the same distribution regardless which of the consistent honest inputs $\vec{x}_H$ were actually used.
+This is conditional independence, and since we assume the simulator is correct (Π is secure),
+we have that independence between the honest secrets and the corrupt real-world views conditional on the corrupt ideal-world views
+is a requisite feature of MPC security.
+In other words
+
+$$
+\left( \vec{y}_H \centernot{⫫} \mathsf{Real}_C(Π, \vec{x}) \mid \mathsf{Ideal}_C(F,\vec{x}) \right)
+\rightarrow
+\nexists \; \mathtt{SIM}
+$$
+
+We use the techniques from \cite{decision trees paper}
+to directly test for conditional dependence without any static analysis of the subject protocol.
 
 To actually test a protocol, one uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
-One of the two models has access to the "real world" views, and the only has access to the "ideal world" views.
-In principal, \textit{any} ML system is ok to use; the choice is partly an engineering question
+One of the two models has access to the real-world views, and the only has access to the ideal-world views.
+In principal, \textit{any} ML classifier is ok to use; the choice is partly an engineering question
 and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
@@
 The process is enumerated in Algorithm 1.
-
-Note that the data must be understood as vectors/matricies/tensors of bits. [is this actually important? are there exceptions/precondidtions?]
+Note that the data must be understood as vectors/matrices/tensors of bits;
+aggregating values into larger structures such as bytes or integers would at best require careful modifications to Algorithm 1.
 
 \begin{algorithm}  %% https://tex.stackexchange.com/a/219820/244649
-  \caption{Does it need a name?}
+  \caption{Test for MPC insecurity}
   \begin{algorithmic}[1]
-  \STATE Set a chosen number of iterations `iters`,
-         number of rows of data on which to train each model `trainN`,
-         and number of rows of data on which to test each model `testN`.
+    \STATE Set a chosen number of iterations \texttt{iters},
+           number of rows of data on which to train each model \texttt{trainN},
+           and number of rows of data on which to test each model \texttt{testN}.
   \STATE Choose $\alpha$, your probability of randomly flagging a secure protocol as insecure.
-  \FOR {`i <- [1..iters]`, calculate scores for the real and ideal models:}
-    \STATE Run the protocol `trainN` times, collecting the corrupt views and honest secrets.
+    \FOR {\texttt{i$\leftarrow$[1..iters]},
+          calculate scores for the real and ideal models:  \\}
+      \STATE Run the protocol \texttt{trainN} times, collecting the corrupt views and honest secrets.
       \STATE Train the first ML model on the collected data.
-           The real-world views are the input features, and the honest secrets are the labels. [I think this is generic ML terminology?]
-    \STATE Train the second ML model the same way, but occlude (omit or overwrite) view data to only reveal the ideal-world view.
-    \STATE Run the protocol `testN` more times, again collecting the relevant data.
+             The real-world views are the input features, and the honest secrets are the labels.
+      \STATE Train the second ML model the same way, but occlude view data to only reveal the ideal-world view.
+      \STATE Run the protocol \texttt{testN} more times, again collecting the relevant data.
       \STATE The score for each model is the $L^1$ distance between the testing labels and the respective model outputs
-           (\ie the number of bits wrong when run on the `testN` rows of test data).
+             (\ie the number of bits wrong when run on the \texttt{testN} rows of test data).
     \ENDFOR
-  \STATE Using the `iters` samples of "real-world" and "ideal-world" scores,
-         calculate the $p$-value for the hypothesis that the distribution of real-world scores is better on average than the ideal-world distribution.
-  \IF {$p \leq \alpha$} \RETURN `INSECURE` \ELSE \RETURN `SECURE` \ENDIF
+    \STATE Using the \texttt{iters} samples of real-world and ideal-world scores,
+           calculate the $p$-value for the hypothesis that
+           the distribution of real-world scores is better on average than the ideal-world distribution.
+    \IF {$p \leq \alpha$} \RETURN \texttt{INSECURE} \ELSE \RETURN \texttt{MAYBE SECURE} \ENDIF
   \end{algorithmic}
@@
 
-\subsection{Issue: false positives}
+\subsection{Theoretical issues}
 
-What our tool actually invalidates is that the views are uniformly random within the slice compatible with the obseved inputs and outputs.
+In principal the only aspect of MPC (in)security that conditional (in)dependence fails to capture
+is the constraint on a hypothetical adversary's computational power.
+In other words, a protocol may truly have the appropriate conditional independence,
+but still be insecure because access to the real-world views allows an adversary to compute in polynomial-time
+something that would require super-polynomial-time in the ideal world.
+We believe this to be a much smaller concern in practice than the statistical limitations of independence testing.
 
-\section{The tool in practice}
+The usual trade-offs of significance testing apply.
+Setting $\alpha$ too low will cause programs who's insecure-ness is only marginally discernible to the D-Trees
+to pass the test.
+Setting $\alpha$ higher will compromise one's certainty that a failing program truly is insecure.
+There are three reasons why this is not a serious problem with the test, each worth considering as its own subtlety.
 
-[How weedy do we want to get here?]
+First, if the test returns \texttt{INSECURE},
+then there's nothing wrong with running the test again to see if you were just unlucky.
+This assumes that you have the resources (\eg compute time) to do so, which you might not.
+Since more powerful tests are better able to detect problems,
+if your compute-budget is fixed you're better off spending it all on a single bigger test.
+
+Second, we observe that the transition of an insecure program between "appears secure" and "appears insecure"
+as the power of the test increases is sharp,
+and as the size of the test continues to increase the $p$-values quickly drop below any reasonable $\alpha$.
+This can be frustrating; if you're unsure of the security of a protocol
+then there's always a chance a modestly-more-powerful test might be conclusive.
+On the other hand, it means that at any reasonable fixed test-size,
+most of the protocols it can detect insecure will be \textit{unambiguously} insecure.
+
+The third such mitigating factor is also a problem on its own.
+For a secure program the observed $p$-values
+will typically \textit{not} be sampled from the uniform distribution on $\mathbb{R}_{(0,1]}$,
+and $\alpha$ is actually an \textit{upper bound} on the likelyhood of falsely flagging a protocol as insecure.
+Contrary to casual descriptions of significance testing, this is not a problem with the statistics;
+it's a result of the inequality in the testing hypothesis.
+[TODO: footnote: Significance testing of an \textit{equality} based null-hypothesis will yield uniformly distributed $p$-values
+when the null-hypothesis is true.]
+Because the ML models have no knowlege of the importance or semantics of the bits they're trained on,
+the mostly-random message- and tape-data provided to the real-world models makes it harder for them
+to learn from the input- and output-data they share with the ideal-world models.
+In other words, the additional data on which the real-world models are trained is fundementally a hindrance to them,
+and only becomes advantageous if the protocol is "insecure enough" to overcome this hidrance.
+A protocol might be truely insecure and not have the relevant conditional independence,
+but still appear secure because the advantage an adversary could gain at guessing the honest inputs
+is smaller than the advantage the ideal-world ML models gain from their more focused data.
+This issue can be overcome by setting ...................
+
+\subsection{Performance analysis}
+
+[TODO: any commentary about the actual python script?]
 
@@
 
-\subsection{Performance analysis}
-
-\section{Case Study Protocols}
+\subsection{Case-study protocols}
 \label{sec:case-study-protocols}

### 2023-11-01 paper/main.tex
describe the tool

@@
           calculate scores for the real and ideal models:  \\}
-      \STATE Run the protocol \texttt{trainN} times, collecting the corrupt views and honest secrets.
+      \STATE Run the protocol \texttt{trainN} times,
+             with uniformly random inputs,
+             collecting the corrupt views and honest secrets.
       \STATE Train the first ML model on the collected data.
@@
 is smaller than the advantage the ideal-world ML models gain from their more focused data.
-This issue can be overcome by setting ...................
+This issue could be overcome by setting \texttt{trainN} sufficiently high;
+the ideal-world models' scores will saturate before the real-world models' do.
 
-\subsection{Performance analysis}
+One might fix the above problem by padding the ideal-world views with random noise out to the same width as the real-world views.
+In our experiments this introduced an even worse problem.
+The additional data afforded to the real-world models is not fully random;
+the whole view is self-consistent per the semantics of the protocol.
+Therefore, the uniformly-random padding given to the ideal-world model will be "heavier" than what it's supposed to be balancing,
+invalidating a key assumption of the whole test:
+that any advantage of the real-world models comes from insecurity in the protocol.
+
+\section{Performance analysis}
 

### 2023-11-01 paper/main.tex
edits

@@
 
-\newcommand{\system}{\textsc{SystemName}\xspace}
+\newcommand{\system}{\textsc{\textbf{.}CHO}\xspace}
 
@@
 
+\subsection{Implementation}
+
+Our python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
+Annotation of which columns are available in the real and ideal worlds is provided in the header.
+The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
+The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
+and they're trained in parallel with \texttt{joblib.Parallel}.
+The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
+and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.
+
+TODO: link to github or other archive?
+
 \section{Performance analysis}
-
-[TODO: any commentary about the actual python script?]
 
@@
 
-Our python script for actually performing the test consumes a stream of fixed-length bit vectors.
-(In practice, this is a CSV. Annotation of which columns are available in the real and ideal worlds is provided in the header.)
-[Does our tool strictly need the vectors to be fixed width? Could someone do this kind of test on a system with branching control flow?]
-To facilitate analysis of the tool, we implemented a choreographic programming language, CHO,
+To facilitate analysis of the tool, we implemented a choreographic programming language, \system,
 with surrounding tooling to support large batches of parallel evaluations of a single program and piping of the generated data.
-The design of CHO is driven by the contextual assumptions of MPC security.
-[Do we want a BNF? Semantics?]
+(In particular, the semantics are polymorphic within the Haskell type-class \texttt{Data.Bits.FiniteBits},
+\eg sixty four evaluations can run in a single thread by applying the semantics to the \texttt{Word64} type.)
+The design of \system is driven by the contextual assumptions of MPC security.
 In addition to bit-level computations and transmission of single bits between arbitrary parties,
-each party has its own read-only tape for secret inputs and random bits
-(CHO is also a probabilistic language),
+each party has its own read-only tapes for secret inputs and randomness,
 and its own tape for output bits.
-CHO supports macros, but they are extremely limited.
-Finally, CHO supports 1-of-N oblivious transfer as a primitive operation.
+\system supports macros, but they are extremely limited.
+Finally, \system supports 1-of-N oblivious transfer as a primitive operation.
 This allows us to test (correct and incorrect) implementations of the classic GMW protocol.
+
+TODO: refer to an example figure and/or an appendix.
 
@@
 Such a tools is only useful if there if new protocols and implementations are being written.
-Therefore, while it is critical to demonstrate that our tool can detect errors manually edite into known-secure protocols,
+Therefore, while it is critical to demonstrate that our tool can detect errors manually edited into known-secure protocols,
 we \textit{also} need to show that the tool can detect a lack of security in entirely novel and maximally diverse protocols.
@@
 The probability that a randomly generated protocol will be MPC secure is not completely negligible.
-As an easy example, if no communicated values ever depend on secret inputs, then the protocol meets the definition of security!
-(Never mind that such a protocol is useless; \textit{none} of the protocols we generated are \textit{useful}.)
-That said, for reasonable generator settings we find that \textit{very few} generated protocols
+In particular, some generator settings will be biased toward vacuous protocols in which nothing is communicated
+or vacuously secure protocols in which the honest secrets are fully revealed by the ideal functionality.
+(Never mind that such protocols are useless; none of the protocols generated in this experiment are \textit{useful}.)
+That said, for reasonable generator settings we find that very few generated protocols
 continue to appear secure as we increase the power of the test.

### 2023-11-02 paper/main.tex
edits
add references

@@
-\documentclass[conference]{IEEEtran}
+\documentclass[compsoc, conference, a4paper, 10pt, times]{IEEEtran}
 
@@
 
-\cite{https://arxiv.org/pdf/1806.07197.pdf} % Computer-aided proofs for multiparty computation with active security
+Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
 use EasyCrypt to manually show the security of GMW.
@@
 
-\cite{Barth et al} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
+Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
 give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
@@
 
-\cite{Gancher et al} % https://dl.acm.org/doi/pdf/10.1145/3571223
+Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
 give a powerful language system that can check if one protocol is a valid simulator of another,
@@
 
-\cite{Wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
-give fine-grain integration of MPC as a language primitive, to help a larger program use it efficiently.
+The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
+give fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
 [TODO: there are other/later works along these lines we should cite?]
 
-\cite{lambda-obliv} %
+Darais et al.~\cite{darais2019language} %
 gives a type system that ensures uniform and independent randomness of key values throughout a program.
@@
 
-\cite{OWL} % https://eprint.iacr.org/2023/473.pdf
+OWL~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
 gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
@@
 
-\cite{Fournet et al} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
+Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
 also use a type system to ensure safe usage of cryptographic primitives,
@@
 
-We experimented with choreography languages that translate directly into Sum Product Networks (SPNs) \cite{TODO}.
+We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
 A SPN is directed-graph representation of a multi-variate distribution;
@@
 
-The \cite{DICE} language analogously translates programs into binary decision diagrams (BDDs).
+The DICE language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
 BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
@@
 
-\cite{lilac} % https://arxiv.org/abs/2304.01339
-is a proof-assistant system for reasoning about probability distributions and PPLs.
+Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
+is a program logic for reasoning about probability distributions and PPLs.
 It more expressive than the original probabilistic separation logic,
@@
 
-\cite{Kifer} % https://arxiv.org/pdf/1805.10277.pdf
-and more recently \cite{Zhang et al} % https://dl.acm.org/doi/10.1145/3428233
+Ding et al.~\cite{ding2018detecting} % https://arxiv.org/pdf/1805.10277.pdf
+and more recently Zhang et al.~\cite{zhang2020testing} % https://dl.acm.org/doi/10.1145/3428233
 use large samples of program executions to test for violations of differential privacy,
-showing that such techniques can work on cryptographic distributional hyper-properties.
+showing that such techniques can work on probabilistic hyper-properties.
 Since differential privacy is only concerned with program inputs and outputs,
@@
 
-Our work builds directly on \cite{decision trees paper} % where did it go? all i can find is https://arxiv.org/pdf/2007.12325.pdf which definitely isn't what we worked off of...
+Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast} % where did it go? all i can find is https://arxiv.org/pdf/2007.12325.pdf which definitely isn't what we worked off of...
 who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
@@
 We attempt to show this by randomly generating hundreds [can we say thousands?] of protocols and testing them with different test parameters.
+We also use a batch of randomly generated protocols to asses the effects of the three test parameters
+(\texttt{iters}, \texttt{trainN}, and \texttt{testN})
+on the test's power and runtime.
 
@@
 
+\bibliographystyle{plain}
+\bibliography{refs}
+
 \end{document}

### 2023-11-02 paper/main.tex
WIP: evaluation section

@@
 
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
 \usepackage{algorithm,algorithmic}  %% typset algorithms
-\usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage{booktabs}
+\usepackage[hidelinks]{hyperref}
+
 \usepackage{amsthm}
-\usepackage{graphicx}
 %%\usepackage{newpxtext}  %% for _e.g._ macro.
@@
 
-\newcommand{\system}{\textsc{\textbf{.}CHO}\xspace}
+\newcommand{\langname}{\textsc{\textbf{.}CHO}\xspace}
+\newcommand{\toolname}{\textsc{DT-SIM}\xspace}
 
@@
 
-\section{Performance analysis}
+\section{Evaluation}
+
+We evaluate the test in terms of three standards of performance:
+\begin{itemize}
+\item \textbf{RQ1:} Is \toolname capable of detecting insecurity in novel protocols outside existing MPC paradigms?
+\item \textbf{RQ2:} Is \toolname capable of detecting insecurity in implementations of real MPC protocols?
+\item \textbf{RQ3:} Is \toolname scalable to realistic protocols in reasonable time?
+\end{itemize}
+We address \textbf{RQ1} in Section~\ref{sec:random-protocols}
+and \textbf{RQ2} in Section~\ref{sec:case-study-protocols}.
+In both cases we measure runtime and discuss implications for \textbf{RQ3}.
+To facilitate all our experiments,
+We built a custom DSL for representing protocols at the appropriate granularity and for easy integration with the test script
+(Section~\ref{sec:cho-lang}).
 
 \subsection{A Choreography Language}
+\label{sec:cho-lang}
 
-To facilitate analysis of the tool, we implemented a choreographic programming language, \system,
+To facilitate analysis of the tool, we implemented a choreographic programming language, \langname,
 with surrounding tooling to support large batches of parallel evaluations of a single program and piping of the generated data.
@@
 \eg sixty four evaluations can run in a single thread by applying the semantics to the \texttt{Word64} type.)
-The design of \system is driven by the contextual assumptions of MPC security.
+The design of \langname is driven by the contextual assumptions of MPC security.
 In addition to bit-level computations and transmission of single bits between arbitrary parties,
@@
 and its own tape for output bits.
-\system supports macros, but they are extremely limited.
-Finally, \system supports 1-of-N oblivious transfer as a primitive operation.
+\langname supports macros, but they are extremely limited.
+Finally, \langname supports 1-of-N oblivious transfer as a primitive operation.
 This allows us to test (correct and incorrect) implementations of the classic GMW protocol.
@@
 \subsection{Random generation of programs}
+\label{sec:random-protocols}
 
@@
 
-[Probably a lot of the details of the generator go in the appendix?]
+\paragraph{Experiment setup.}
+In our first experiment, we generate random protocols and evaluate \toolname's ability to detect insecurity in these protocols.
+We randomly generate protocols according to the grammar of \langname.
+The generator uses a specified size for each party's secret inputs, random tape, and outputs,
+parameters for the length of the "body" of the program and the "max width" of each computation step,
+and a handful of bias-parameters that affect the frequency of different operations.
+We bias the generated protocols in favor of logical depth;
+values (inputs, coin-flips, messages-received, values calculated) are likely
+to be used on each new line in inverse proportion to the number of times they've already been used.
+We generate XXXX programs of size XXXX and evaluate \toolname's ability to detect insecurity.
+
+\paragraph{Results.}
+\emph{results description goes here}
 
@@
 
-For our evaluation, we implemented six MPC protocols from the literature in \system, and evaluated \system's scalability and ability to find implementation bugs in these protocols.
-The six protocols are listed in Table~\ref{tbl:protocols}.
+For our evaluation, we implemented three boolean functions, each in two variations of the GMW protocol,
+and evaluated \toolname's scalability and ability to find bugs added to these protocols.
+The function-circuits and protocols are listed in Table~\ref{tbl:protocols}.
 One is a simple addition protocol, three are basic building blocks often used to build more complex protocols, and two are protocols designed to securely evaluate circuits.
@@
 To ensure security, the input values are \emph{secret shared} using additive secret sharing; to secret share a bit, a party selects a random bit to represent one share, and computes the other share by XORing the random bit with the secret bit.
-In \system, $P_1$ can secret share the bit $x$ with $P_2$ as follows:
+In \langname, $P_1$ can secret share the bit $x$ with $P_2$ as follows:
 %
@@
 
-Our $n$-bit addition protocol generates a \system source file for a given bitwidth.
-In our evaluation, we test several different bitwidths to examine the effect of the size of each party's view on the scalability and bug-finding capability of \system.
+Our $n$-bit addition protocol generates a \langname source file for a given bitwidth.
+In our evaluation, we test several different bitwidths to examine the effect of the size of each party's view on the scalability and bug-finding capability of \toolname.
 
@@
 Beaver triples can be generated by a trusted ``dealer'' party, or by using oblivious transfer.
-We implement both versions in \system.
+We implement both versions in \langname.
 
@@
 The parties evaluate multiplication gates by using 1-out-of-4 oblivious transfer: the first party generates a random bit for its share of the output, and acts as the sender in OT; the second party acts as the receiver in OT and receives its share of the output.
-We implemented a compiler to generate \system programs that evaluate  Bristol fashion circuits using GMW, and evaluated it on circuits implementing integer addition and integer inequalities.
+We implemented a compiler to generate \langname programs that evaluate  Bristol fashion circuits using GMW, and evaluated it on circuits implementing integer addition and integer inequalities.
 
@@
 Our implementation of this approach employees a third-party ``dealer'' who generates the triples; a pre-processing (offline) phase can also be used to generate the triples.
-We evaluate \system on this protocol using the same circuits as in our GMW experiments.
+We evaluate \toolname on this protocol using the same circuits as in our GMW experiments.
 
-% \paragraph{Circuits.}
-% To evaluate the circuit-based protocols, we used five binary circuits that are publicly available in the ``Bristol-fashion'' format~\cite{todo}.
-We chose commonly-used circuits with a range of sizes to test \system's scalability.
-Table~\ref{tbl:circuits} lists the circuits used in our evaluation, along with the number of gates in each circuit.
-
-% \begin{table}
-% \centering
-%   \begin{tabular}{|l |l|l|}
-%     \hline
-%     \textbf{Circuit} & \textbf{AND Gates} & \textbf{Total Gates} \\
-%     \hline
-%     64-bit adder & 63 & 376 \\
-%     64-bit multiplier & 4,033 & 13,675 \\
-%     AES-128 & 6,400 & 36,663 \\
-%     SHA-256 & 22,573 & 135,073 \\
-%     Floating-point division & 82,269 & 184,007 \\
-%     \hline
-%   \end{tabular}
-%   \caption{Circuits used in our evaluation.}
-%   \label{tbl:circuits}
-% \end{table}
-
-\section{Evaluation}
-
-Our empirical evaluation demonstrates \system's ability to detect bugs in secure protocols. Our experiments are designed to answer two research questions:
-%
-\begin{itemize}
-\item \textbf{RQ1:} Is \system capable of detecting insecurity in complex, multi-round protocols?
-\item \textbf{RQ2:} Is \system capable of detecting insecurity in real-world protocols?
-\item \textbf{RQ3:} Is \system scalable to realistic protocols in reasonable time?
-\end{itemize}
-%
-To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \system's ability to show they are insecure.
+Our empirical evaluation demonstrates \toolname's ability to detect bugs in secure protocols.
+To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \toolname's ability to show they are insecure.
 The details appear in Section~\ref{sec:experiment-1:-random}.
-To answer RQ2, we evaluate \system on the case study protocols described in Section~\ref{sec:case-study-protocols}.
+To answer RQ2, we evaluate \toolname on the case study protocols described in Section~\ref{sec:case-study-protocols}.
 The details appear in Section~\ref{sec:experiment-2:-real}.
-In both cases, \system is able to detect insecurity in the target protocols in under a minute.
-
-\subsection{Experiment 1: Random Protocols}
-\label{sec:experiment-1:-random}
+In both cases, \toolname is able to detect insecurity in the target protocols in under a minute.
 
 \paragraph{Experiment setup.}
-In our first experiment, we generate random protocols and evaluate \system's ability to detect insecurity in these protocols.
-We randomly generate protocols according to the grammar of XXX.
-We bias the generated protocols in favor of generating and using randomness in each round of the protocol, in addition to the secret inputs.
-Random protocols of non-trivial length are extremely likely to be insecure.
-%
-We generate XXXX programs of size XXXX and evaluate \system's ability to detect insecurity.
-
-\paragraph{Results.}
-\emph{results description goes here}
-
-\subsection{Experiment 2: Real-world Protocols}
-\label{sec:experiment-2:-real}
-
-\paragraph{Experiment setup.}
-In our second experiment, we evaluate \system's ability to detect insecurity in the real-world protocols described in Section~\ref{sec:case-study-protocols}.
-We run \system on the unmodified protocol in order to establish that the protocol is secure.
-Then, we mutate the protocol in several ways to introduce insecurity, and test \system's ability to detect it.
+In our second experiment, we evaluate \toolname's ability to detect insecurity in the real-world protocols described in Section~\ref{sec:case-study-protocols}.
+We run \toolname on the unmodified protocol in order to establish that the protocol is secure.
+Then, we mutate the protocol in several ways to introduce insecurity, and test \toolname's ability to detect it.
 Specifically, our mutations are:

### 2023-11-02 paper/main.tex
add initial graphs

@@
 \usepackage{algorithm,algorithmic}  %% typset algorithms
+
 \usepackage{graphicx}
@@
 % use quotes as smart-quotes. https://tex.stackexchange.com/a/52354
-\usepackage [english]{babel}
+%\usepackage [english]{babel}
 \usepackage [autostyle, english = american]{csquotes}
 \MakeOuterQuote{"}
+\usepackage{graphicx}
 
@@
 
+\appendices
+
+\section{Additional Experimental Results}
+
+% figures for the main body
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.45\textwidth}
+\begin{tabular}{c| c c}
+    \hline\hline
+  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
+    \hline\hline
+  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
+  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_128_1024.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
+  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_128_1024.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
+  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_128_1024.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results.}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.45\textwidth}
+\begin{tabular}{c| c c}
+    \hline\hline
+  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
+    \hline\hline
+  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
+  & \includegraphics[width=\gsize]{graphs/time_adder_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_adder_beaver_256_2048.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
+  & \includegraphics[width=\gsize]{graphs/time_less_than_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_less_than_beaver_256_2048.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
+  & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_beaver_256_2048.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: decision tree training time.}
+\end{figure*}
+
+\input{full_figures.tex}
+
 \end{document}

### 2023-11-02 paper/full_figures.tex
add full figures tex

% These are full figures for the appendix

\begin{figure*}
  \centering
  \newcommand{\gsize}{.45\textwidth}
\begin{tabular}{c| c c}
    \hline\hline
  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
    \hline\hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_128_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_128_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_128_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_128_2048.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_256_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_256_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_256_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_256_2048.pdf} \\
    \hline
    \hline
\end{tabular}
\caption{Experimental results, $n$-bit addition.}
\end{figure*}

\begin{figure*}
  \centering
  \newcommand{\gsize}{.45\textwidth}
\begin{tabular}{c| c c}
    \hline\hline
  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
    \hline\hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_128_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_128_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_128_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_128_2048.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_256_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_256_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_256_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_256_2048.pdf} \\
    \hline
    \hline
\end{tabular}
\caption{Experimental results, $n$-bit less-than comparison.}
\end{figure*}

\begin{figure*}
  \centering
  \newcommand{\gsize}{.45\textwidth}
\begin{tabular}{c| c c}
    \hline\hline
  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
    \hline\hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_128_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_128_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 128, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_128_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_128_2048.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 1024$}
  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_256_1024.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_256_1024.pdf} \\
    \hline
  \rotatebox{90}{\phantom{helloh}$i = 256, n = 2048$}
  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_256_2048.pdf}
                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_256_2048.pdf} \\
    \hline
    \hline
\end{tabular}
\caption{Experimental results, $n$-bit Beaver triple generation.}
\end{figure*}

### 2023-11-02 paper/main.tex
move graphs into evaluation section
add random-cho graphs

@@
 
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.9\textwidth}
+\begin{tabular}{c | c}
+    \hline\hline
+    \rotatebox{90}{(a). Detection of insecurity}
+  & \includegraphics[width=\gsize]{graphs/medium2.old.pdf} \\
+    \hline
+    \rotatebox{90}{(b). Test time}
+  & \includegraphics[width=\gsize]{graphs/asymptote_time.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: performance on random protocols under increasing test power.}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.45\textwidth}
+\begin{tabular}{c | c c}
+    \hline\hline
+    \rotatebox{90}{(a). Detection scores}
+  & \includegraphics[width=\gsize]{graphs/cube_scores_back.pdf}
+  & \includegraphics[width=\gsize]{graphs/cube_scores_front.pdf} \\
+    \hline
+    \rotatebox{90}{(b). Test time}
+  & \includegraphics[width=\gsize]{graphs/cube_times_back.pdf}
+  & \includegraphics[width=\gsize]{graphs/cube_times_front.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: trade-offs in power and runtime under different test parameters.}
+\end{figure*}
+
 \subsection{Case-study protocols}
@@
 
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.45\textwidth}
+\begin{tabular}{c| c c}
+    \hline\hline
+  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
+    \hline\hline
+  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
+  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_128_1024.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
+  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_128_1024.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
+  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_128_1024.pdf}
+                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_128_1024.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: ability to detect security bugs.}
+\end{figure*}
+
 \paragraph{Oblivious Transfer.}
@@
 
+\begin{figure*}
+  \centering
+  \newcommand{\gsize}{.45\textwidth}
+\begin{tabular}{c| c c}
+    \hline\hline
+  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
+    \hline\hline
+  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
+  & \includegraphics[width=\gsize]{graphs/time_adder_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_adder_beaver_256_2048.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
+  & \includegraphics[width=\gsize]{graphs/time_less_than_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_less_than_beaver_256_2048.pdf} \\
+    \hline
+  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
+  & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_gmw_256_2048.pdf}
+                 & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_beaver_256_2048.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: decision tree training time.}
+\end{figure*}
+
 \paragraph{Experiment setup.}
@@
 
-% figures for the main body
-\begin{figure*}
-  \centering
-  \newcommand{\gsize}{.45\textwidth}
-\begin{tabular}{c| c c}
-    \hline\hline
-  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
-    \hline\hline
-  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
-  & \includegraphics[width=\gsize]{graphs/security_adder_gmw_128_1024.pdf}
-                 & \includegraphics[width=\gsize]{graphs/security_adder_beaver_128_1024.pdf} \\
-    \hline
-  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
-  & \includegraphics[width=\gsize]{graphs/security_less_than_gmw_128_1024.pdf}
-                 & \includegraphics[width=\gsize]{graphs/security_less_than_beaver_128_1024.pdf} \\
-    \hline
-  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
-  & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_gmw_128_1024.pdf}
-                 & \includegraphics[width=\gsize]{graphs/security_beaver_triple_gen_beaver_128_1024.pdf} \\
-    \hline
-    \hline
-\end{tabular}
-\caption{Experimental results.}
-\end{figure*}
-
-\begin{figure*}
-  \centering
-  \newcommand{\gsize}{.45\textwidth}
-\begin{tabular}{c| c c}
-    \hline\hline
-  & \textbf{OT-Based (GMW)} & \textbf{Beaver Triple-Based}\\
-    \hline\hline
-  \rotatebox{90}{\phantom{hellohello}$n$-bit addition}
-  & \includegraphics[width=\gsize]{graphs/time_adder_gmw_256_2048.pdf}
-                 & \includegraphics[width=\gsize]{graphs/time_adder_beaver_256_2048.pdf} \\
-    \hline
-  \rotatebox{90}{\phantom{hel}$n$-bit less-than comparison}
-  & \includegraphics[width=\gsize]{graphs/time_less_than_gmw_256_2048.pdf}
-                 & \includegraphics[width=\gsize]{graphs/time_less_than_beaver_256_2048.pdf} \\
-    \hline
-  \rotatebox{90}{\phantom{h}$n$-bit Beaver triple generation}
-  & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_gmw_256_2048.pdf}
-                 & \includegraphics[width=\gsize]{graphs/time_beaver_triple_gen_beaver_256_2048.pdf} \\
-    \hline
-    \hline
-\end{tabular}
-\caption{Experimental results: decision tree training time.}
-\end{figure*}
-
 \input{full_figures.tex}

### 2023-11-02 paper/main.tex
updating evaluation section part 2

@@
 
+\paragraph{Oblivious Transfer.}
+Oblivious transfer~\cite{todo} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
+In 1-out-of-2 OT, $S$ provides 2 secret inputs $x_0$ and $x_1$ and $R$ provides a secret \emph{selection bit} $b$.
+If $b=0$, $R$ receives $x_0$; if $b=1$, $R$ receives $x_1$.
+$R$ does not learn anything about the not-selected secret, and $S$ does not learn which secret $R$ selected.
+This idea can be extended to 1-out-of-$n$ OT by extending the selection bit $b$ to a $\log_2(n)$-bit string.
+%
+Oblivious transfer is a building block for many other protocols.
+
 TODO: refer to an example figure and/or an appendix.
@@
 
-\subsection{Case-study protocols}
+\subsection{Bug-Finding in Real-World Protocols}
 \label{sec:case-study-protocols}
 
-For our evaluation, we implemented three boolean functions, each in two variations of the GMW protocol,
-and evaluated \toolname's scalability and ability to find bugs added to these protocols.
-The function-circuits and protocols are listed in Table~\ref{tbl:protocols}.
-One is a simple addition protocol, three are basic building blocks often used to build more complex protocols, and two are protocols designed to securely evaluate circuits.
+Our second set of experiments evaluates \toolname's ability to detect security bugs in real-world protocols.
+We implement two MPC frameworks for evaluating circuits, instantiate them with three different functions to obtain specialized protocols, systematically introduce bugs into these protocols, and evaluate \toolname's ability to detect the bugs.
+We summarize the implemented protocols in Section~\ref{sec:protocols-evaluated}, describe the experiment in Section~\ref{sec:e2-experiment-setup}, and present the results in Section~\ref{sec:e2_results}.
+
+\subsubsection{Protocols Evaluated}
+\label{sec:protocols-evaluated}
+
+For our evaluation, we consider two MPC frameworks for evaluating circuits, and instantiate those frameworks with three different functions (represented as circuits) to construct secure protocols.
+The frameworks and functions are listed in Table~\ref{tbl:protocols} and described below.
 All six protocols are designed for two parties, with the addition of a third ``dealer'' party to produce correlated randomness in several of the protocols.
@@
   \centering
-  \begin{tabular}{|l | l|l|}
+  \renewcommand*{\arraystretch}{1.2}
+  \begin{tabular}{|l l|}
     \hline
-    \textbf{Protocol} & \textbf{Purpose} & \textbf{Ref.} \\
+    \textbf{MPC Frameworks} & \textbf{Approach} \\
     \hline
-    $n$-bit Addition & Custom & TODO \\
-    Oblivious Transfer & Building block & TODO \\
+    %Oblivious Transfer & Building block & TODO \\
 %    OT Extension & Building block & TODO \\
-    Beaver Triple Generation & Building block & TODO \\
-    GMW & Circuit evaluation & TODO \\
-    MPC w/ Beaver Triples & Circuit evaluation & TODO \\
+    GMW~\cite{goldreich2019play, goldreich2009foundations} & Oblivious Transfer\\
+    MPC w/ Beaver Triples~\cite{beaver1992efficient} & Beaver Triples\\[4pt]
+    \hline
+    \textbf{Instantiated Functions} & \textbf{Leakage}\\
+    \hline
+    $n$-bit Addition & All \\
+    $n$-bit Less-than Comparison & Partial \\
+    Beaver Triple Generation & None \\
     \hline
   \end{tabular}
-  \caption{Secure protocols used in our evaluation.}
+  \bigskip
+  \caption{MPC frameworks and instantiated functions used in our evaluation.
+Each framework describes an approach for evaluating a binary circuit in MPC; we instantiate both frameworks with three different functions (represented as circuits) to construct secure protocols.
+}
   \label{tbl:protocols}
@@
 
-\paragraph{$n$-bit Addition.}
-The simplest protocol in our evaluation is a basic summation protocol, which adds up the inputs from the two participating parties.
+\paragraph{GMW (Goldreich-Micali-Widgerson).}
+The GMW protocol~\cite{goldreich2019play, goldreich2009foundations} is an MPC protocol for evaluating circuits.
+Each party holds \emph{additive secret shares} of each wire's value; to secret share a bit, a party selects a random bit to represent one share, and computes the other share by XORing the random bit with the secret bit.
+%
+The parties evaluate addition gates (and other linear operations) by adding the shares they hold.
+%
+The parties evaluate multiplication gates by using 1-out-of-4 oblivious transfer: the first party generates a random bit for its share of the output, and acts as the sender in OT; the second party acts as the receiver in OT and receives its share of the output.
+We implemented a compiler to generate \langname programs that evaluate Bristol fashion circuits using GMW.
+
+\paragraph{MPC with Beaver Triples.}
+Another MPC protocol for evaluating circuits also uses additive shares to hold wire values, but uses \emph{Beaver triples}~\cite{beaver1992efficient} to evaluate multiplication gates (instead of OT).
+A Beaver triple (or \emph{multiplication triple}) is a set of secret shares of numbers $a$, $b$, and $c$ such that $a \cdot b = c$.
+To use a Beaver triple, $P_1$ must have one share of each of $a$, $b$, and $c$, and $P_2$ must have the other share.
+Since the values are secret shared, they appear random to each party---but the correlation between the three numbers can be used to perform efficient multiplication between secret-shared numbers (described later).
+One Beaver triple must be consumed for each multiplication, but they can be generated in a separate offline phase.
+
+Given shares of a Beaver triple, the parties perform two multiplication-by-constant and broadcast operations, and derive one share each of the product.
+Our implementation of this approach employees a third-party ``dealer'' who generates the triples; a pre-processing (offline) phase can also be used to generate the triples.
+We implemented a similar compiler to generate \langname programs leveraging this approach for circuit specifications.
+
+\paragraph{Functions evaluated.}
+We evaluate three functions, which we express as binary circuits for two parties and compile to \langname programs using the approach described above.
+All three functions expect input from both the honest and corrupt party.
+The first, an $n$-bit addition circuit, is vacuously secure---its ideal functionality leaks all of the honest party's inputs, because the corrupt party can subtract their inputs from the $n$-bit result to reconstruct the honest party's inputs.
+The second, an $n$-bit less-than comparison circuit, only leaks a single bit of information---which of the two inputs is larger.
+For $n=1$, the less-than protocol is also vacuously secure, but it leaks only partial information about the honest party's inputs for $n>1$.
+The third, an $n$-bit Beaver triple generator, leaks no information about the honest party's input, since its output to each party is a single secret share of the Beaver triple.
+
+% \paragraph{$n$-bit Addition.}
+% The simplest protocol in our evaluation is a basic summation protocol, which adds up the inputs from the two participating parties.
 To accomplish this task, the protocol essentially simulates a hardware adder circuit---for each bit in the binary representation of the two input numbers, the protocol performs bitwise operations to compute the sum and carry bit.
@@
 In \langname, $P_1$ can secret share the bit $x$ with $P_2$ as follows:
-%
-\begin{verbatim}
-  s1 = FLIP @P1
-  s2 = x + s1
-  SEND s1 TO P2
-\end{verbatim}
-%
-$P_1$ can safely send one share to $P_2$, as long as $P_1$ keeps the other share for themselves---neither share alone reveals anything about the secret bit.
+% %
+% \begin{verbatim}
+%   s1 = FLIP @P1
+%   s2 = x + s1
+%   SEND s1 TO P2
+% \end{verbatim}
+% %
+% $P_1$ can safely send one share to $P_2$, as long as $P_1$ keeps the other share for themselves---neither share alone reveals anything about the secret bit.
 To add (i.e.
@@
 
-Our $n$-bit addition protocol generates a \langname source file for a given bitwidth.
+% Our $n$-bit addition protocol generates a \langname source file for a given bitwidth.
 In our evaluation, we test several different bitwidths to examine the effect of the size of each party's view on the scalability and bug-finding capability of \toolname.
@@
 \end{tabular}
-\caption{Experimental results: ability to detect security bugs.}
+\caption{Experimental results: ability to detect security bugs.
+We set the number of iterations to 128 and the training set size to 1024.
+Results for other settings appear in Appendix~\ref{sec:addit-exper-results}.}
+\label{fig:results_security}
 \end{figure*}
-
-\paragraph{Oblivious Transfer.}
-Oblivious transfer~\cite{todo} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
-In 1-out-of-2 OT, $S$ provides 2 secret inputs $x_0$ and $x_1$ and $R$ provides a secret \emph{selection bit} $b$.
-If $b=0$, $R$ receives $x_0$; if $b=1$, $R$ receives $x_1$.
-$R$ does not learn anything about the not-selected secret, and $S$ does not learn which secret $R$ selected.
-This idea can be extended to 1-out-of-$n$ OT by extending the selection bit $b$ to a $\log_2(n)$-bit string.
-%
-Oblivious transfer is a building block for many other protocols.
 
@@
 
-\paragraph{Beaver Triple Generation.}
-A \emph{Beaver triple}~\cite{todo} (or \emph{multiplication triple}) is a set of secret shares of numbers $a$, $b$, and $c$ such that $a \cdot b = c$.
-To use a Beaver triple, $P_1$ must have one share of each of $a$, $b$, and $c$, and $P_2$ must have the other share.
-Since the values are secret shared, they appear random to each party---but the correlation between the three numbers can be used to perform efficient multiplication between secret-shared numbers (described later).
-One Beaver triple must be consumed for each multiplication, but they can be generated in a separate offline phase.
-Beaver triples can be generated by a trusted ``dealer'' party, or by using oblivious transfer.
-We implement both versions in \langname.
-
-\paragraph{GMW (Goldreich-Micali-Widgerson).}
-The GMW protocol is an MPC protocol for evaluating circuits.
-Each party holds additive secret shares of each wire's value, and evaluates addition gates by adding the shares they hold.
-The parties evaluate multiplication gates by using 1-out-of-4 oblivious transfer: the first party generates a random bit for its share of the output, and acts as the sender in OT; the second party acts as the receiver in OT and receives its share of the output.
-We implemented a compiler to generate \langname programs that evaluate  Bristol fashion circuits using GMW, and evaluated it on circuits implementing integer addition and integer inequalities.
-
-\paragraph{MPC with Beaver Triples.}
-Another MPC protocol for evaluating circuits also uses additive shares to hold wire values, but uses Beaver triples to evaluate multiplication gates (instead of OT).
-Given shares of a Beaver triple, the parties perform two multiplication-by-constant and broadcast operations, and derive one share each of the product.
-Our implementation of this approach employees a third-party ``dealer'' who generates the triples; a pre-processing (offline) phase can also be used to generate the triples.
-We evaluate \toolname on this protocol using the same circuits as in our GMW experiments.
-
-Our empirical evaluation demonstrates \toolname's ability to detect bugs in secure protocols.
-To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \toolname's ability to show they are insecure.
+% Our empirical evaluation demonstrates \toolname's ability to detect bugs in secure protocols.
+% To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \toolname's ability to show they are insecure.
 The details appear in Section~\ref{sec:experiment-1:-random}.
@@
 \caption{Experimental results: decision tree training time.}
+\label{fig:results_training_time}
 \end{figure*}
 
-\paragraph{Experiment setup.}
-In our second experiment, we evaluate \toolname's ability to detect insecurity in the real-world protocols described in Section~\ref{sec:case-study-protocols}.
-We run \toolname on the unmodified protocol in order to establish that the protocol is secure.
-Then, we mutate the protocol in several ways to introduce insecurity, and test \toolname's ability to detect it.
+\subsubsection{Experiment Setup}
+\label{sec:e2-experiment-setup}
+
+We run \toolname on the unmodified protocol in order to establish the protocol's security.
+Then, we mutate the protocol in several ways to introduce security bugs, and test \toolname's ability to detect these bugs.
 Specifically, our mutations are:
@@
 
-\paragraph{Results.} \emph{results description goes here}
+\subsubsection{Results}
+\label{sec:e2_results}
+
+The results appear in Figures~\ref{fig:results_security} and~\ref{fig:results_training_time}.
+For the results in Figure~\ref{fig:results_security}, we set the number of iterations to 128 and the training set size to 1024.
+The results are similar for other settings; additional experimental results appear in Appendix~\ref{sec:addit-exper-results}.
+
+\paragraph{Detecting security bugs.}
+For the $n$-bit addition function,...
+
+For the $n$-bit less-than comparison,...
+
+For the $n$-bit Beaver triple generation function,...
+
+\paragraph{Scalability.}
+The primary computational cost of \toolname is the time taken to train the decision trees.
+We record the total training time for each combination of MPC framework and function, and plot the results for several settings in Figure~\ref{fig:results_training_time}.
+For the settings we evaluated, training time grows linearly with the number of iterations, the training data size, and the bitwidth of the target function (i.e.
+the size of the circuit being tested).
+Even for the largest bitwidths we considered, training takes only seconds to minutes.
 
@@
 \section{Additional Experimental Results}
+\label{sec:addit-exper-results}
 

### 2023-11-02 paper/main.tex
tweaked graphs and evaluation section

@@
 \begin{itemize}
+\item \textbf{Biased sharing}: uses biased randomness to construct secret shares of honest party inputs.
+\item \textbf{Biased AND}: uses biased randomness as part of the evaluation of multiplication gates.
 \item \textbf{Accidental secret}: accidentally sends a secret held by the honest party  to the corrupt party.
 \item \textbf{Accidental gate}: accidentally sends an honest party's result from the evaluation of a multiplication gate to the corrupt party.
-\item \textbf{Biased sharing}: uses biased randomness to construct secret shares of honest party inputs.
-\item \textbf{Biased AND}: uses biased randomness as part of the evaluation of multiplication gates.
 \end{itemize}
@@
 The severity of each mutation can be tuned by adjusting the bias in the related randomness.
-In the case of ``accidental secret'' and ``accidental gate,'' this affects the probability of accidentally revealing a value; in the others, it directly adjust the bias in the randomness.
+In the case of \textbf{accidental secret} and \textbf{accidental gate}, this affects the probability of accidentally revealing a value; in the others, it directly adjust the bias in the randomness.
 
@@
 \paragraph{Detecting security bugs.}
-For the $n$-bit addition function,...
+\toolname is able to detect most security bugs in our evaluation, but interpreting its results requires care.
+\toolname detects \textbf{biased sharing} bugs in nearly all cases: the associated $p$-values are universally close to 0 in all six cases.
+The one exception is the $n$-bit adder for small $n$, where associated $p$ values are close to 1.
+This is likely due to the \emph{vacuous security} of the $n$-bit adder: for small bitwidths, the decision trees are able to reconstruct the honest party's inputs using only information from the ideal functionality, rendering \emph{any} protocol secure.
+For larger bitwidths, the decision trees are no longer able to perform this reconstruction, and the protocols are classified as insecure.
+The other functions---which are not vacuously secure---do not display this behavior.
 
-For the $n$-bit less-than comparison,...
+\toolname detects \textbf{biased AND} bugs in most of the GMW-based protocols, but struggles to do so in the case of Beaver-triple-based protocols.
+Due to the use of multiplication triples, taking advantage of the leakage from biased AND gates in this case is much more difficult than in the GMW case.
+These results suggest that \toolname will likely have difficulty detecting bugs that require significant ``backwards computation'' to leverage the leaked information.
 
-For the $n$-bit Beaver triple generation function,...
+\toolname detects \textbf{accidental secret} bugs in the same cases as biased sharing bugs, for the same reasons.
+
+\toolname detects \textbf{accidental gate} bugs only in the $n$-bit less-than comparison based on the GMW framework.
+In the Beaver-triple framework, we hypothesize that the decision trees are not capable of the ``backwards computation'' required to leverage the leaked information; in the other cases, the function's design means that AND gate outputs do not provide additional leakage beyond the ideal functionality.
+
+\toolname fails to detect \textbf{biased AND} and \textbf{accidental gate} bugs in the case of $n$-bit Beaver triple generation because the result of each AND gate is also one of the protocol's outputs---so a bug which reveals it does not leak any new information.
+\toolname detects \textbf{accidental gate} bugs for the $n$-bit Beaver triple generation function when the Beaver-triple-based protocol is used, because biasing the randomness for AND gates reveals information about the Beaver triples themselves, from which it is possible to deduce the honest party's shares of the inputs.
 
@@
 \subsection{Discussion}
+
+The results of this experiment suggest that \toolname scales linearly with protocol size and test power, and is capable of detecting bugs in real-world protocols.
+The test typically runs in just seconds to minutes.
+The results of our experiment on protocol mutations show that \toolname is effective for detecting important classes of security bugs, but that care must be taken in interpreting the results.
+
+For security bugs that directly reveal sensitive information (e.g.
+\textbf{biased sharing} bugs) \toolname is particularly effective.
+Our results suggest that \toolname is capable of detecting these accidental disclosures even at low test powers and in the context of large protocols.
+
+\toolname is less effective for security bugs that require significant ``backwards computation'' to leverage the leaked information (e.g.
+\textbf{accidental gate} bugs), since decision trees are not well-suited to solving this problem.
+A more sophisticated model---such as a deep neural network---may be able to perform this kind of computation, but would take significantly longer to train.
 

### 2023-11-02 paper/main.tex
changes to graphs

--- 2023-11-02 paper/main.tex
@@
 \paragraph{Experiment setup.}
-In our first experiment, we generate random protocols and evaluate \toolname's ability to detect insecurity in these protocols.
+In our first two experiments, we generate random protocols and evaluate \toolname's ability to detect insecurity in these protocols.
 We randomly generate protocols according to the grammar of \langname.
@@
 to be used on each new line in inverse proportion to the number of times they've already been used.
-We generate XXXX programs of size XXXX and evaluate \toolname's ability to detect insecurity.
+Finally, to focus our experiment on protocols for which there's any question of security,
+\langname protocols that \toolname flags as insecure with very low power-settings
+are filtered out of the stream of generated programs before we include them in our experiments.
 
-\paragraph{Results.}
-\emph{results description goes here}
+We generated 500 programs with 500-line bodies (16-bit inputs and outputs, 48-bit random tapes)
+and evaluate \toolname's ability to detect insecurity as \texttt{trainN} increased.
+Within this experiment we locked \texttt{iters} at 100, and set \texttt{testN} as close to $\frac{\mathtt{trainN}}{4}$ as our setup allowed.
+Each generated \langname file was tested with increasing values of \texttt{trainN} until the observed $p$-values fell below $1.25 \cdot 10^{-4}$.
+
+To observe the relative role of the three test parameters, we generated 200 additional programs using the same settings
+and ran \toolname against each of them with a grid of different parameters.
 
 \begin{figure*}
+  \label{fig:sankey}
   \centering
   \newcommand{\gsize}{.9\textwidth}
-\begin{tabular}{c | c}
+\begin{tabular}{c}
     \hline\hline
-    \rotatebox{90}{(a). Detection of insecurity}
-  & \includegraphics[width=\gsize]{graphs/medium2.old.pdf} \\
-    \hline
-    \rotatebox{90}{(b). Test time}
-  & \includegraphics[width=\gsize]{graphs/asymptote_time.pdf} \\
+    \includegraphics[width=\gsize]{graphs/medium2.old.pdf} \\
     \hline
@@
 
+\begin{figure}
+  \centering
+  \newcommand{\gsize}{.38\textwidth}
+\begin{tabular}{c}
+    \hline\hline
+    \includegraphics[width=\gsize]{graphs/asymptote_time.pdf} \\
+    \hline
+    \hline
+\end{tabular}
+\caption{Experimental results: runtime with increasing test power.}
+\end{figure}
+
 \begin{figure*}
@@
     \hline\hline
-    \rotatebox{90}{(a). Detection scores}
+    \rotatebox{90}{\phantom{hellohello}(a). Detection scores (summed)}
   & \includegraphics[width=\gsize]{graphs/cube_scores_back.pdf}
@@
     \hline
-    \rotatebox{90}{(b). Test time}
+    \rotatebox{90}{\phantom{hellohellohello}(b). Runtime (s)}
   & \includegraphics[width=\gsize]{graphs/cube_times_back.pdf}
@@
 \end{figure*}
+
+\paragraph{Results.}
+Figure~\ref{fig:sankey} shows that, while some programs re...
 

### 2023-11-02 paper/main.tex
Random-CHOs:Results

@@
 \begin{figure}
+  \label{fig:linear-time}
   \centering
@@
 \begin{figure*}
+  \label{fig:cube-search}
   \centering
@@
     \hline
-    \rotatebox{90}{\phantom{hellohellohello}(b). Runtime (s)}
+    \rotatebox{90}{\phantom{hellohellohello}(b). Mean runtime (s)}
   & \includegraphics[width=\gsize]{graphs/cube_times_back.pdf}
@@
 \paragraph{Results.}
-Figure~\ref{fig:sankey} shows that, while some programs re...
+Figure~\ref{fig:sankey} shows that, while some programs appear secure at weaker test powers,
+all the programs generated with these settings were identified as insecure with $\texttt{iters}=100$ and $\texttt{trainN}\leq4000$.
+Specifically, this Sankey plot shows 500 randomly generated \langname files moving between buckets labeled
+"\texttt{\#\_sec}" (for secure),
+"\texttt{\#\_ins}" (for insecure at $\alpha=0.05$),
+and "\texttt{\#\_neg}" (for "negligible", insecure at $\alpha=1.25 \cdot 10^{-4}$)
+as they are tested by \toolname with increasing \texttt{trainN}.
+Because there's always a chance (at most $\alpha$) of a secure protocol being flagged insecure,
+\langname{}s sometimes move back to "secure" from "insecure"; we continue testing them until they reach "negligible".
+Figure~\ref{fig:linear-time} shows time data from this same experiment;
+because fewer programs were subjected to the larger tests, the minimum and maximum runtimes converge at the end.
+More importantly, the runtime scales linearly with the amount of data used in the test.
+
+Figure~\ref{fig:cube-search} shows the individual effects of \texttt{iters}, \texttt{trainN}, and \texttt{testN} on \toolname's power and runtime.
+Each 3D heatmap shows aggregate data for 200 randomly generated \toolname files.
+Because is mostly driven by the amount of data consumed ($\texttt{iters} \cdot (\texttt{trainN} + \texttt{testN})$),
+all three affect runtime linearly.
+To maximize the granularity of the test-power analysis, we calculate a score for each combination of parameters,
+where apparently-secure programs count for 2 points,
+programs insecure at $\alpha=0.05$ count for 1 point,
+and programs insecure at $\alpha=0.01$ count for zero.
+Lower scores are "better" because we know all the programs turn out to be actually insecure.
+We observe that scaling all three parameters is desirable,
+but \texttt{trainN} should generally be larger than \texttt{testN},
+and \texttt{iters} can be much smaller.
 

### 2023-11-02 paper/main.tex
intro tweaks and some more refs

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{natbib}
 
@@
 
-\title{Title Goes Here}
+\title{\toolname: Property-Based Testing for MPC Security}
 
-\author{First1 Last1}
+%\author{First1 Last1}
 %%\authornote{with author1 note}        %% and a great more deleted...
@@
   but such methods are still labor intensive and the need remains for an automated tool that can positively identify an insecure protocol.
-  We show that the philosophy of property-based-testing, \textit{"run it a bunch of times and see if it breaks"},
-  is at least somewhat feasible in a cryptographic setting.
-  We specifically target Secure Multi-Party Computation, because formal methods targeting this security definition for bit-model implementations
+  In this work, we demonstrate that property-based testing, \textit{"run it a bunch of times and see if it breaks"},
+  is effective for detecting security bugs in secure protocols.
+  We specifically target Secure Multi-Party Computation (MPC), because formal methods targeting this security definition for bit-model implementations
   are particularly difficult.
@@
 
-\dots  %% TODO: intorduction per se?
+%\subsection{Motivation}
 
-\subsection{Motivation}
-
-As in all kinds of engineering; in cryptography it is critical to know that the thing you are building actually does what you intend.
-In cryptography however, the intended behavior is often expressed as a negation,
-\eg secret information should \textit{never} be derivable from such'n'such surface of the system.
-Observing of the system's behavior provides limited evidence that the system is actually secure in the intended ways.
-Formal proofs of security are a popular and well-suited solution to this problem, however:
-
+Secure multiparty computation (MPC) protocols~\cite{evans2018pragmatic} allow a group of mutually distrusting parties to compute a shared function over distributed data without revealing that data.
+The security of MPC protocols is often expressed as a negation,
+\eg secret information should \textit{never} be derivable from an adversary's observations of the system.
+Unfortunately, testing generally provides limited evidence that the system is actually secure in the intended ways; formal proofs of security are effective, but are tedious to construct and require significant expertise.
+As a result:
+%
 \begin{itemize}
-    \item As consumers, we are concerned that a great deal of cyber-security software has been and is being produced without formal verification.
-    \item As engineers, we would like to not waste time attempting to prove the security of systems which, unbeknownst to us, are not actually secure.
-    \item As product managers, we would like to have sanity checks
-          that complicated software systems actually have the security properties their constituent source-code claims to be proven to have.
+\item Most implementations of MPC protocols are not formally verified.
+\item Formal verification tools are challenging to use, and engineers can waste significant time trying to prove properties of insecure implementations.
+\item Even for formally-verified software, we would like to have sanity checks
+  that protocol implementations actually have the security properties their constituent source-code claims to be proven to have.
 \end{itemize}
 
-In this work we particularly focus on Secure Multi-Party Computation (MPC),
-a cryptographic primitive with wide-ranging applications but frustrating performance limitations.
-Substantial work has going into improving the performance of MPC systems [TODO: citation].
-Additionally, in some cases performance can be improved by using a bespoke protocol instead of a general-purpose MPC system [do we have any evidence of this?].
-We accelerate this work by providing a tool, comparable to a property-based test,
-for conveniently spot-checking the security of an MPC implementation.
+We present \toolname, an \emph{automatic}, \emph{property-based testing} tool for MPC security.
+\toolname works by running the protocol thousands of times, then using a statistical independence test to detect security bugs.
+Our approach is motivated by the success of property-based testing (via tools like \textsc{QuickCheck}) as an approach for finding bugs in software~\cite{fink1997property, claessen2000quickcheck, paraskevopoulou2015foundational}.
+Unlike previous applications of property-based testing, however, \toolname attempts to falsify a \emph{probabilistic hyperproperty} of the target protocol---a much more challenging setting.
 
-\subsection{Contributions}
+\toolname works by defining security in terms of conditional independence, then performing a statistical test which attempts to falsify this conditional independence property.
+In MPC protocols, the inputs of the honest parties should be independent of the \emph{views} (i.e.
+the observed message values) of the corrupt parties, conditioned on the protocol's output.
+Conditioning on the protocol's output is vital for the security definition, since it allows the protocol to \emph{intentionally} leak some information about the inputs via the output---as nearly all MPC protocols do.
+Manual security proofs for MPC protocols demonstrate this conditional independence property by constructing a \emph{simulator} that ``fakes'' the corrupt parties' views using only the protocol's output.
 
-In this work we present and evaluate a statistical test of the security of MPC protocols.
-First we show how MPC security relates to inference and equivalence of probabilistic distributions,
-and then we demonstrate that existing tests of distributional equivalence can, with minor modification, be used to refute MPC security.
-Our tool is adaptable to the wide range of cryptographic security properties that can be defined in terms of distributional equivalence,
-and our tool is applicable to any software system that
-A) can be run 10k-1M times without great inconvenience, and
-B) allows certain intermediate values (of which consumed randomness is the least convenient) to be captured for analysis.
+This formulation of security as conditional independence puts MPC security out of reach of all automated verification tools (and many manual ones).
+While automated tools have been used to verify security in cryptographic protocols~\cite{gancher2023owl, barthe2019probabilistic, darais2019language, fournet2011information}  (e.g.
+key exchange protocols), these tools are limited to protocols whose \emph{outputs reveal nothing} to the adversary.
+They are capable of showing independence, but not the kind of \emph{conditional} independence required for MPC security.
+Recent manual approaches could be used to verify MPC security~\cite{li2023lilac, gancher2023core, haagh2018computer}, but require manual proof.
+
+\toolname is designed to fill the same role as tools like \textsc{QuickCheck}: it can be used to quickly and automatically detect many classes of bugs in MPC protocol implementations, without requiring the programmer to spend time building a manual security proof.
+We have implemented \toolname and tested it empirically on several real-world MPC protocols and on hundreds of randomly-generated protocols; our experimental results suggest that \toolname is capable of scaling to realistic protocols and detecting many security bugs, despite the challenging setting.
+
+% In this work we particularly focus on Secure Multi-Party Computation (MPC),
+% a cryptographic primitive with wide-ranging applications but frustrating performance limitations.
+% Substantial work has going into improving the performance of MPC systems [TODO: citation].
+% Additionally, in some cases performance can be improved by using a bespoke protocol instead of a general-purpose MPC system [do we have any evidence of this?].
+% We accelerate this work by providing a tool, comparable to a property-based test,
+% for conveniently spot-checking the security of an MPC implementation.
+
+\paragraph{Contributions.}
+In summary, our contributions are:
+%
+\begin{itemize}
+\item We show how to define MPC security in terms of conditional independence, enabling the use of independence testing tools for checking MPC security.
+\item We introduce \toolname, the first property-based testing tool for MPC security.
+\item We conduct an empirical evaluation of \toolname, demonstrating its ability to scale to realistic MPC protocols and detect security bugs.
+\end{itemize}
+
+% In this work we present and evaluate a statistical test of the security of MPC protocols.
+% First we show how MPC security relates to inference and equivalence of probabilistic distributions,
+% and then we demonstrate that existing tests of distributional equivalence can, with minor modification, be used to refute MPC security.
+% Our tool is adaptable to the wide range of cryptographic security properties that can be defined in terms of distributional equivalence,
+% and our tool is applicable to any software system that
+% A) can be run 10k-1M times without great inconvenience, and
+% B) allows certain intermediate values (of which consumed randomness is the least convenient) to be captured for analysis.
 
@@
 Sometimes this is done with a select handful of inputs or configurations,
-but non-deterministic programs are better evaluated with \termOfArt{property-based testing},
+but non-deterministic programs are better evaluated with \termOfArt{property-based testing}~\cite{fink1997property, claessen2000quickcheck, paraskevopoulou2015foundational},
 in which the program is run with many arbitrary (but not necessarily random) configurations.
 Property-based testing is a popular paradigm for testing code in industry;
-the classic Haskell framework QuickCheck [citation] % https://www.cse.chalmers.se/~rjmh/QuickCheck/
+the classic Haskell framework \textsc{QuickCheck}~\cite{claessen2000quickcheck}
 is still in use and has been ported to various other languages.

### 2023-11-02 paper/main.tex
WIP: future work

--- 2023-11-02 paper/main.tex
@@
 
-The results of this experiment suggest that \toolname scales linearly with protocol size and test power, and is capable of detecting bugs in real-world protocols.
+The results of these experiments suggest that \toolname scales linearly with protocol size and test power, and is capable of detecting bugs in real-world protocols.
 The test typically runs in just seconds to minutes.
@@
 \section{Future Work}
+
+Some further works is needed to turn \toolname into a deploy-able tool;
+our experiments so far are not sufficient to calculate sensible default settings,
+and we believe some changes to the way the data is ingested and handled by the tool could improve efficiency.
+
+There are also variations on the core concept of \toolname that we believe are worth exploring.
+Decision trees are not the only candidate for the underlying ML model;
+the additional compute-cost of \eg a neural network may turn out to be efficient.
+Other cryptographic properties besides MPC can likely
 

### 2023-11-02 paper/main.tex
conclusions

@@
 the additional compute-cost of \eg a neural network may turn out to be efficient.
-Other cryptographic properties besides MPC can likely
+Other cryptographic properties besides MPC can likely be framed in terms of conditional independence;
+and so would be amenable to similar tests.
+Finally, because end-to-end static analysis of MPC protocols has proven so difficult,
+we believe it may be practical to augment a stochastic tool like \toolname with static analysis capabilities
+that would allow it to more efficiently detect insecurity, turning it into a "white-box" test.
 
 \section{Conclusions}
+
+We present \toolname,
+a property-based test that checks (a small over-approximation of) MPC security
+using a probabilistic test of conditional independence.
+At its heart, \toolname trains batches of decision trees with and without access to grey-box data from many evaluations of the target protocol
+to see if access to that message data improves their ability to predict the secret inputs.
+While this approach has theoretical limitations, we show that in practice it is able to detect various mistakes in various protocols,
+making it a useful tool for sanity-checking one's work during development.
+We also show that, in practice, the compute-time required for \toolname scales linearly with the power of the test requested,
+and usefully-powerful tests take just a few minutes.
 

### 2023-11-02 paper/main.tex
knockinng out TODOs

@@
 
-%\subsection{Motivation}
-
 Secure multiparty computation (MPC) protocols~\cite{evans2018pragmatic} allow a group of mutually distrusting parties to compute a shared function over distributed data without revealing that data.
 The security of MPC protocols is often expressed as a negation,
-\eg secret information should \textit{never} be derivable from an adversary's observations of the system.
+\eg{} secret information should \textit{never} be derivable from an adversary's observations of the system.
 Unfortunately, testing generally provides limited evidence that the system is actually secure in the intended ways; formal proofs of security are effective, but are tedious to construct and require significant expertise.
@@
 \toolname is designed to fill the same role as tools like \textsc{QuickCheck}: it can be used to quickly and automatically detect many classes of bugs in MPC protocol implementations, without requiring the programmer to spend time building a manual security proof.
+Using \toolname requires only that key runtime values can be captured during program execution;
+it's agnostic to the language and many other details about the subject program.
 We have implemented \toolname and tested it empirically on several real-world MPC protocols and on hundreds of randomly-generated protocols; our experimental results suggest that \toolname is capable of scaling to realistic protocols and detecting many security bugs, despite the challenging setting.
-
-% In this work we particularly focus on Secure Multi-Party Computation (MPC),
-% a cryptographic primitive with wide-ranging applications but frustrating performance limitations.
-% Substantial work has going into improving the performance of MPC systems [TODO: citation].
-% Additionally, in some cases performance can be improved by using a bespoke protocol instead of a general-purpose MPC system [do we have any evidence of this?].
-% We accelerate this work by providing a tool, comparable to a property-based test,
-% for conveniently spot-checking the security of an MPC implementation.
 
@@
 Implementations that work on arbitrary functions (representable as circuit computations in some field)
-have existed since 1982 [TODO: citation], % https://ieeexplore.ieee.org/document/4568207
+have existed since 1986~\cite{yao86}, % https://ieeexplore.ieee.org/document/4568207
 but their use by the public is limited by their performance.
@@
 Π results in all parties getting the same outputs as they would have in the ideal world.
-Defining the \textit{security} of Π requires a clear threat model, and requires defining the participants' "views".
+The \textit{security} of Π is expressed in terms of a particular threat model and the participants' "\termOfArt{views}".
 
-\termOfArt{Passive Security} supposes a computationally-bounded adversary
+\paragraph{Passive Security}
+supposes a computationally-bounded adversary
 that can see everything happening inside a fixed subset of the participating machines.
@@
 The \termOfArt{ideal world view}, $\mathsf{Ideal}_P(F,\vec{x})$, of a party P is what it sees when running F in the ideal world:
-just its inputs $\vec{x}_P$ and its outputs.
+just its inputs $\vec{x}_P$ and its outputs $\vec{y}_P$.
 The \termOfArt{real world view}, $\mathsf{Real}_P(Π, \vec{x})$, of a party P is everything it sees while executing its part of Π in the real world:
 its inputs, outputs, all messages it receives, and all the bits of random tape it uses.
-Both these concepts extend to naturally to lists of parties.
+Both these concepts extend naturally to lists of parties.
 
-\begin{definition}[Passive MPC Security]
+\begin{definition}[Passive MPC Security]\label{def:passive-mpc}
     A protocol Π,
@@
 
-This definition represents security by a transitive argument:
+Definition~\ref{def:passive-mpc} represents security by a transitive argument:
 Since the adversary in the ideal world can compute all the data the real-world adversary starts with,
@@
 
-Definition 1 is tight, in the sense that attempting to relax it invariable either
+Definition~\ref{def:passive-mpc} is tight, in the sense that attempting to relax it invariable either
 allows a "secure" protocol to do something bad, or requires introducing restrictions
-on the kinds of functionality to which the definition may be applied.
-(Of course the idea that a secure protocol \textit{can't} do anything bad implies an assumption that computing
-the functionality is a good thing. [move to footnote])
+on the kinds of functionality to which the definition may be applied\footnote{
+    The idea that a secure protocol \textit{can't} do anything bad assumes that computing the functionality is good to begin with.
+}.
 That said, some simplification is possible for our purposes:
@@
 (which are captured in the real-world views),
-we can safely ignore any messages sent \textit{within} C.
-Then we can collapse all the corrupted parties C into a single party C,
-and similarly all the other parties into a single party H (for "honest")
-(we don't always do this).
-Finally, in this work we assume that protocols implement their functionalities \textit{correctly},
-and usually leave the corresponding F implicit.
+we can safely ignore any messages sent \textit{within} $C$.
+Then we can collapse all the corrupted parties $C$ into a single party "C",
+and similarly all the other parties into a single party "H" (for "honest")\footnote{
+    We don't always do this, it's not required for \toolname to work.
+}.
+Finally, in this work we assume that protocols implement their functionalities \textit{correctly};
+the corresponding F is implicit.
 In other words, we define F to be whatever mapping results from the execution of Π.
@@
 
-\termOfArt{Active Security} (sometimes "malicious security")
-supposes a computationally bounded adversary that can cause corrupted machines to take arbitrary actions,
+\paragraph{Active Security}
+(sometimes "malicious security")
+supposes a computationally bounded adversary that can cause corrupted machines to take arbitrary actions:
 breaking the choreography,
@@
 
-Certain edge cases are helpful to consider, as they really do map out the edges of the problem of testing from MPC security.
+\paragraph{Vacuousness.}\footnote{
+    This terminology is new; we needed it to talk though problems we encountered in during this project.
+}
+Certain edge cases are helpful to consider, as they map out the edges of the problem of testing from MPC security.
 In the first case, Π itself may be vacuous.
@@
 
-TODO: maybe say we invented these; give examples. Readers will assume these are existing terms if we don't say otherwise.
-
+\paragraph{Non-interference.}
 While MPC can be expressed as a kind of non-interference property, this is only somewhat useful.
 The non-interference can only be meant within the space of possible program traces \textit{posterior to the output}.
-Similarly, trying to avoid such subtleties in the non-interference property by providing a correctly constrained \texttt{reveal} operator
+Trying to avoid such subtleties in the non-interference property by providing a correctly constrained \texttt{reveal} operator
 isn't promising because constraints on \texttt{reveal} would need to recapitulate the original question of security
@@
 1) Random values generated (by any party) at runtime have known distributions
-(\eg the python expression \texttt{secrets.SystemRandom().random()}
+(\eg{} the python expression \texttt{secrets.SystemRandom().random()}
 will sample at runtime from the best floating-point approximation the machine can provide
@@
 
-Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast} % where did it go? all i can find is https://arxiv.org/pdf/2007.12325.pdf which definitely isn't what we worked off of...
+Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
 who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
@@
 
-We use the techniques from \cite{decision trees paper}
+We use the techniques from \cite{chalupka2018fast}
 to directly test for conditional dependence without any static analysis of the subject protocol.
@@
 then there's nothing wrong with running the test again to see if you were just unlucky.
-This assumes that you have the resources (\eg compute time) to do so, which you might not.
+This assumes that you have the resources (\eg{} compute time) to do so, which you might not.
 Since more powerful tests are better able to detect problems,
@@
 (In particular, the semantics are polymorphic within the Haskell type-class \texttt{Data.Bits.FiniteBits},
-\eg sixty four evaluations can run in a single thread by applying the semantics to the \texttt{Word64} type.)
+\eg{} sixty four evaluations can run in a single thread by applying the semantics to the \texttt{Word64} type.)
 The design of \langname is driven by the contextual assumptions of MPC security.
@@
 Decision trees are not the only candidate for the underlying ML model;
-the additional compute-cost of \eg a neural network may turn out to be efficient.
+the additional compute-cost of \eg{} a neural network may turn out to be efficient.
 Other cryptographic properties besides MPC can likely be framed in terms of conditional independence;

### 2023-11-02 paper/main.tex
knockinng out TODOs

@@
 
-Efficient inversion of programs and fragments of programs turns out to be so important for determining if a protocol satisfies MPC security.
+Efficient inversion of programs and fragments of programs turns out to be important for determining if a protocol satisfies MPC security.
 Since MPC protocols are probabilistic, what we really need is efficient posterior inference of program inputs in a
@@
 For application to the subject of MPC protocols, a PPL would need to handle the two different kinds of uncertainty facing a hypothetical adversary:
-1) Random values generated (by any party) at runtime have known distributions
-(\eg{} the python expression \texttt{secrets.SystemRandom().random()}
+1) Random values generated (by any party) at runtime have known distributions\footnote{
+    \eg{} in python \texttt{secrets.SystemRandom().random()}
     will sample at runtime from the best floating-point approximation the machine can provide
-of the uniform distribution on $\mathbb{R}_{(0,1]}$).
+    of the uniform distribution on $\mathbb{R}_{(0,1]}$
+}.
 2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
@@
 B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
-of the prior distribution of honest secrets[TODO: footnote], and
-C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs, presumably by handling them symbolically.
+of the prior distribution of honest secrets\footnote{
+    Presumably such process would need to leave those parameters implicit.
+    Attempting to represent the adversary's prior beliefs more compactly would not produce a valid proof or test of security.
+}, and
+C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs.
 Such a system is not necessarily impossible, but we have so far failed to build it.
-
-TODO: Is efficient inference really necessary? The conditional independence test might be sufficient? I think we don't want to imply that inference is sufficient, because lots of work has done that
 

### 2023-11-02 paper/main.tex
background & related work move

@@
 \newcommand{\toolname}{\textsc{DT-SIM}\xspace}
+
+\newcommand{\mynote}[2]
+    {{\color{red} \fbox{\bfseries\sffamily\scriptsize#1}
+    {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}}~}
+%\newcommand{\mynote}[2]{}
+\newcommand{\todo}[1]{\mynote{TODO}{#1}}
 
@@
 Implementations that work on arbitrary functions (representable as circuit computations in some field)
-have existed since 1986~\cite{yao86}, % https://ieeexplore.ieee.org/document/4568207
+have existed since 1986~\cite{yao1986generate}, % https://ieeexplore.ieee.org/document/4568207
 but their use by the public is limited by their performance.
@@
 
-The traditional expression of MPC assumes a pre-determined function F, "the functionality",
+The traditional expression of MPC assumes a pre-determined function F, the \emph{functionality},
 mapping the tuple of all participants' input vectors to the respective tuple of output vectors
@@
 
-\paragraph{Passive Security}
+\textbf{Passive security}
 supposes a computationally-bounded adversary
@@
 
+\todo{this is the first place the simulator is mentioned, but it's not described}
 \begin{definition}[Passive MPC Security]\label{def:passive-mpc}
@@
 
-\paragraph{Active Security}
+\textbf{Active security}
 (sometimes "malicious security")
@@
 breaking the choreography,
-forging messages (to such extent as that's actually possible),
+forging messages,
 lying about computation results,
@@
 \paragraph{Non-interference.}
-While MPC can be expressed as a kind of non-interference property, this is only somewhat useful.
+While MPC can be expressed as a kind of probabilistic non-interference property~\cite{haagh2018computer, almeida2018enforcing}, this is only somewhat useful.
 The non-interference can only be meant within the space of possible program traces \textit{posterior to the output}.
-Trying to avoid such subtleties in the non-interference property by providing a correctly constrained \texttt{reveal} operator
-isn't promising because constraints on \texttt{reveal} would need to recapitulate the original question of security
+Trying to avoid such subtleties in the non-interference property by providing a correctly constrained declassification operator
+does not avoid this problem: constraints on declassification would need to recapitulate the original question of security
 and examine the entire rest of the protocol as context.
-
-TODO: maybe state it explicitly with the conditioning? Then we can say that no system can capture that property.
 
@@
 
-Efficient inversion of programs and fragments of programs turns out to be important for determining if a protocol satisfies MPC security.
-Since MPC protocols are probabilistic, what we really need is efficient posterior inference of program inputs in a
-\termOfArt{probabilistic programming language} (PPL).
-
+Since MPC security can be framed as a test of conditional independence, \termOfArt{probabilistic programming languages} (PPLs) with efficient inference and independence tests could be used to verify MPC security.
+%
 A PPL is a language that provides stochastic non-deterministic behavior as a language primitive
@@
 For application to the subject of MPC protocols, a PPL would need to handle the two different kinds of uncertainty facing a hypothetical adversary:
-1) Random values generated (by any party) at runtime have known distributions\footnote{
-    \eg{} in python \texttt{secrets.SystemRandom().random()}
+(1) Random values generated (by any party) at runtime have known distributions\footnote{
+    \eg{} in Python \texttt{secrets.SystemRandom().random()}
     will sample at runtime from the best floating-point approximation the machine can provide
     of the uniform distribution on $\mathbb{R}_{(0,1]}$
-}.
-2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
+}, and
+(2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
 possible secrets representing their prior belief so that the secret inputs can also be modeled as random variables.
@@
 \textit{if and only if} the inferred distribution were independent of the adversary's choice of prior distribution for the honest secrets.
-So a direct test of MPC security (or a useful over-approximation) could be built from
-A) a language meta-theory guarantee that such posterior inference and sampling would be efficient,
-B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
+
+A direct test of MPC security via a PPL would require \emph{both}
+(A) a language meta-theory guarantee that such posterior inference and sampling would be efficient, \emph{and}
+(B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
 of the prior distribution of honest secrets\footnote{
@@
     Attempting to represent the adversary's prior beliefs more compactly would not produce a valid proof or test of security.
-}, and
-C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs.
-Such a system is not necessarily impossible, but we have so far failed to build it.
+}, as well as
+(C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs, presumably by handling them symbolically.
+While some PPLs can statically verify independence properties~\cite{gorinova2021conditional}, none can verify the kind of \emph{conditional} independence required for MPC security.
+Lilac~\cite{li2023lilac} provides a separation logic for reasoning about conditional independence, but is not automated.
 
@@
 the various used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
-
-\section{Related Work}
-
-\subsection{Formal-methods for cryptography}
-Although a useful language-driven test for MPC security remains out of reach,
-a variety of useful work aproaches the subject from different directions.
-
-Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
-use EasyCrypt to manually show the security of GMW.
-Their definitions generalize to protocols with efficiently invertible functionalities.
-% remove? Since existing general-purpose SMPC protocols can easily compute non-invertible functions, this is a serious limitation.
-They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
-If their system could be automated and paired with language constraints to ensure invertibility then it
-could apply to many, but certainly not all, use cases.
-
-Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
-give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
-Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.
-
-Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
-give a powerful language system that can check if one protocol is a valid simulator of another,
-and show that it works on existing MPC protocols like GMW.
-The user of the system needs to write a second protocol
-representing the composition of the ideal-world functionality and the explicit function $\mathtt{SIM}$,
-and then build a high-level proof of the equivalence in the provided equational logic.
-
-The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
-give fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
-[TODO: there are other/later works along these lines we should cite?]
-
-Darais et al.~\cite{darais2019language} %
-gives a type system that ensures uniform and independent randomness of key values throughout a program.
-Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
-but this would of course prevent any non-vacuous functionality from being implemented.
-
-OWL~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
-gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
-They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
-there's no way to verify an implementation of MPC.
-
-Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
-also use a type system to ensure safe usage of cryptographic primitives,
-but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
-Like OWL, their type system assumes a secure implementation of the encryption scheme.
-
-\subsection{PPLs for Inference}
-
-We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
-A SPN is directed-graph representation of a multi-variate distribution;
-they are compact, allow efficient inference, and can represent values symbolically as needed,
-so testing for conditional independence is practical.
-However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
-because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.
-
-The DICE language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
-BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
-but since the graph \textit{structure} of the BDD of a DICE program depends on the actual \textit{values} on which one is conditioning,
-a dice-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
-
-Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
-is a program logic for reasoning about probability distributions and PPLs.
-It more expressive than the original probabilistic separation logic,
-but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
-
-\subsection{Probabilistic testing}
-
-Ding et al.~\cite{ding2018detecting} % https://arxiv.org/pdf/1805.10277.pdf
-and more recently Zhang et al.~\cite{zhang2020testing} % https://dl.acm.org/doi/10.1145/3428233
-use large samples of program executions to test for violations of differential privacy,
-showing that such techniques can work on probabilistic hyper-properties.
-Since differential privacy is only concerned with program inputs and outputs,
-the dimensionality of the problem is somewhat smaller than MPC.
-
-Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
-who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
-Specifically, if decision trees (D-Trees trained on random variables $A$ and $C$ do better at predicting the value of a random variable $B$,
-then $A$ and $B$ are not independent conditioned on $C$.
-The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
-In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
-but in our adaptation we use fresh program traces for everything.
 
@@
 
-Our python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
+Our Python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
 Annotation of which columns are available in the real and ideal worlds is provided in the header.
@@
 
+\section{Related Work}
+
+\subsection{Formal-methods for cryptography}
+Although automated language-based verification of MPC security is an open question,
+a variety of useful work aproaches the subject from different directions.
+
+Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
+use EasyCrypt to manually show the security of GMW.
+Their definitions generalize to protocols with efficiently invertible functionalities.
+% remove? Since existing general-purpose SMPC protocols can easily compute non-invertible functions, this is a serious limitation.
+They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
+If their system could be automated and paired with language constraints to ensure invertibility then it
+could apply to many, but certainly not all, use cases.
+
+Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
+give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
+Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.
+
+Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
+give a powerful language system that can check if one protocol is a valid simulator of another,
+and show that it works on existing MPC protocols like GMW.
+The user of the system needs to write a second protocol
+representing the composition of the ideal-world functionality and the explicit function $\mathtt{SIM}$,
+and then build a high-level proof of the equivalence in the provided equational logic.
+
+The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
+give fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
+[TODO: there are other/later works along these lines we should cite?]
+
+Darais et al.~\cite{darais2019language} %
+gives a type system that ensures uniform and independent randomness of key values throughout a program.
+Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
+but this would of course prevent any non-vacuous functionality from being implemented.
+
+OWL~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
+gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
+They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
+there's no way to verify an implementation of MPC.
+
+Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
+also use a type system to ensure safe usage of cryptographic primitives,
+but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
+Like OWL, their type system assumes a secure implementation of the encryption scheme.
+
+\subsection{PPLs for Inference}
+
+We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
+A SPN is directed-graph representation of a multi-variate distribution;
+they are compact, allow efficient inference, and can represent values symbolically as needed,
+so testing for conditional independence is practical.
+However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
+because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.
+
+The DICE language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
+BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
+but since the graph \textit{structure} of the BDD of a DICE program depends on the actual \textit{values} on which one is conditioning,
+a dice-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
+
+Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
+is a program logic for reasoning about probability distributions and PPLs.
+It more expressive than the original probabilistic separation logic,
+but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
+
+\subsection{Probabilistic testing}
+
+Ding et al.~\cite{ding2018detecting} % https://arxiv.org/pdf/1805.10277.pdf
+and more recently Zhang et al.~\cite{zhang2020testing} % https://dl.acm.org/doi/10.1145/3428233
+use large samples of program executions to test for violations of differential privacy,
+showing that such techniques can work on probabilistic hyper-properties.
+Since differential privacy is only concerned with program inputs and outputs,
+the dimensionality of the problem is somewhat smaller than MPC.
+
+Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
+who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
+Specifically, if decision trees (D-Trees trained on random variables $A$ and $C$ do better at predicting the value of a random variable $B$,
+then $A$ and $B$ are not independent conditioned on $C$.
+The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
+In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
+but in our adaptation we use fresh program traces for everything.
+
 \section{Future Work}

### 2023-11-03 paper/main.tex
tweaking related work

@@
 They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
-If their system could be automated and paired with language constraints to ensure invertibility then it
-could apply to many, but certainly not all, use cases.
+Their approach requires manual proof in EasyCrypt.
 
@@
 give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
-Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.
+However, the framework is not amenable to automation.
 
@@
 gives a type system that ensures uniform and independent randomness of key values throughout a program.
-Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
-but this would of course prevent any non-vacuous functionality from being implemented.
+Their system is automated, and could be adapted to ensure perfect noninterference between honest secrets and corrupt views, but does not support the kind of \emph{conditional} independence required for most MPC protocols.
 
@@
 gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
-They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
-there's no way to verify an implementation of MPC.
+OWL is designed for secure protocols that reveal nothing to the adversary, and does not support verification of MPC security.
 
@@
 but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
-Like OWL, their type system assumes a secure implementation of the encryption scheme.
+Like OWL, their type system assumes a secure implementation of the encryption scheme and does not support verification of MPC security.
 
@@
 
-We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
-A SPN is directed-graph representation of a multi-variate distribution;
-they are compact, allow efficient inference, and can represent values symbolically as needed,
+Several PPLs directly support efficient inference and some forms of independence testing.
+A Sum Product Network (SPN)~\cite{poon2011sum} is a directed-graph representation of a multi-variate distribution; SPPL~\cite{saad2021sppl} is a PPL that performs inference using SPNs.
+SPNs are compact, allow efficient inference, and can represent values symbolically as needed,
 so testing for conditional independence is practical.
@@
 It more expressive than the original probabilistic separation logic,
-but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
+but does not support automation.
 
@@
 There are also variations on the core concept of \toolname that we believe are worth exploring.
-Decision trees are not the only candidate for the underlying ML model;
-the additional compute-cost of \eg{} a neural network may turn out to be efficient.
+Decision trees are not the only candidate for the underlying ML model, and our evaluation suggests that they are incapable of performing some of the ``backwards computation'' needed to demonstrate insecurity;
+a more capable model (\eg{} a neural network) may be more capable of detecting these kinds of security bugs.
 Other cryptographic properties besides MPC can likely be framed in terms of conditional independence;

### 2023-11-03 paper/main.tex
knockinng out TODOs

--- 2023-11-02 paper/main.tex
@@
 
-At face value, property-based testing doesn't generally extend to hyper-properties,
+At face value, property-based testing doesn't extend well to hyper-properties,
 which can only be assessed by comparing multiple program traces.
-In practice sometimes one can express a hyper-property as a property of multiple of duplicate programs,
+In practice sometimes one can express a hyper-property as a property of multiple duplicate programs,
 but the sensitivity of the test quickly decays and the logistics of setting up the test quickly grow.
 Furthermore, when the property to be tested is over distributions of traces,
-the various used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
+the techniques used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
 
@@
 Their definitions generalize to protocols with efficiently invertible functionalities.
-% remove? Since existing general-purpose SMPC protocols can easily compute non-invertible functions, this is a serious limitation.
-They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
-If their system could be automated and paired with language constraints to ensure invertibility then it
-could apply to many, but certainly not all, use cases.
+They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
+If their system could be automated and paired with language constraints to ensure invertibility,
+then it could apply to many, but certainly not all, use cases.
 
@@
 The user of the system needs to write a second protocol
-representing the composition of the ideal-world functionality and the explicit function $\mathtt{SIM}$,
+representing the composition of the ideal-world functionality and an explicit $\mathtt{SIM}$,
 and then build a high-level proof of the equivalence in the provided equational logic.
@@
 The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
-give fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
-[TODO: there are other/later works along these lines we should cite?]
+gives fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
+Their system assumes (at least one) secure implementation of MPC on circuits.
 
-Darais et al.~\cite{darais2019language} %
+$\lambda_\textbf{obliv}$~\cite{darais2019language} %
 gives a type system that ensures uniform and independent randomness of key values throughout a program.
@@
 
-OWL~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
+\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
 gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
-They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
+They can type-check convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
 there's no way to verify an implementation of MPC.
@@
 also use a type system to ensure safe usage of cryptographic primitives,
-but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
+but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
 Like OWL, their type system assumes a secure implementation of the encryption scheme.
@@
 We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
-A SPN is directed-graph representation of a multi-variate distribution;
-they are compact, allow efficient inference, and can represent values symbolically as needed,
+A SPN is a directed-graph representation of a multi-variate distribution.
+They are compact, allow efficient inference, and can represent values symbolically as needed,
 so testing for conditional independence is practical.
@@
 
-The DICE language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
+The \textsc{Dice} language~\cite{holtzen2020scaling} translates programs into binary decision diagrams (BDDs),
+analogous to the way we tried to translate them into SPNs.
 BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
-but since the graph \textit{structure} of the BDD of a DICE program depends on the actual \textit{values} on which one is conditioning,
-a dice-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
+but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
+a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
 
@@
 is a program logic for reasoning about probability distributions and PPLs.
-It more expressive than the original probabilistic separation logic,
+It more expressive than the original probabilistic separation logic~\cite{barthe2019probabilistic},
 but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
@@
 who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
-Specifically, if decision trees (D-Trees trained on random variables $A$ and $C$ do better at predicting the value of a random variable $B$,
+Specifically, if decision trees (D-Trees) trained on random variables $A$ and $C$
+do better at predicting the value of a random variable $B$ than D-Trees trained only on $C$,
 then $A$ and $B$ are not independent conditioned on $C$.
@@
 
-We use the techniques from \cite{chalupka2018fast}
-to directly test for conditional dependence without any static analysis of the subject protocol.
+We use the techniques from \cite{chalupka2018fast} to build \toolname,
+which directly tests for the above conditional dependence without any static analysis of the subject protocol.
 
-To actually test a protocol, one uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
+To actually test a protocol, \toolname uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
 One of the two models has access to the real-world views, and the only has access to the ideal-world views.
@@
 Note that the data must be understood as vectors/matrices/tensors of bits;
-aggregating values into larger structures such as bytes or integers would at best require careful modifications to Algorithm 1.
+aggregating values into larger structures such as bytes or integers would at best require careful modifications to line 8 in Algorithm 1.
 
@@
 Setting $\alpha$ higher will compromise one's certainty that a failing program truly is insecure.
-There are three reasons why this is not a serious problem with the test, each worth considering as its own subtlety.
+There are three reasons why this is not a serious problem with \toolname, each worth considering as its own subtlety.
 
-First, if the test returns \texttt{INSECURE},
-then there's nothing wrong with running the test again to see if you were just unlucky.
+First, if \toolname returns \texttt{INSECURE},
+then there's nothing wrong with running it again to see if you were just unlucky.
 This assumes that you have the resources (\eg{} compute time) to do so, which you might not.
@@
 On the other hand, it means that at any reasonable fixed test-size,
-most of the protocols it can detect insecure will be \textit{unambiguously} insecure.
+most of the protocols \toolname \textit{can} detect insecure will be \textit{unambiguously} insecure.
 
@@
 Contrary to casual descriptions of significance testing, this is not a problem with the statistics;
-it's a result of the inequality in the testing hypothesis.
-[TODO: footnote: Significance testing of an \textit{equality} based null-hypothesis will yield uniformly distributed $p$-values
-when the null-hypothesis is true.]
+it's a result of the inequality in the testing hypothesis.\footnote{
+    Significance testing of an \textit{equality} based null-hypothesis would yield uniformly distributed $p$-values when the null-hypothesis is true.
+}
 Because the ML models have no knowlege of the importance or semantics of the bits they're trained on,
-the mostly-random message- and tape-data provided to the real-world models makes it harder for them
+the (mostly random) message- and tape-data provided to the real-world models makes it harder for them
 to learn from the input- and output-data they share with the ideal-world models.
@@
 
-Our python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
+Our \toolname python script consumes a stream of fixed-length bit vectors as a CSV.
 Annotation of which columns are available in the real and ideal worlds is provided in the header.
@@
 
-TODO: link to github or other archive?
+See Section~\ref{sec:public-repo} for more details.
 
@@
 
+\section*{Data Availability}\label{sec:public-repo}
+
+TODO.
+
+This section, and everything following, does not count within the body text page limit requirement (so this paper would satisfy a page limit of 1).
+For your Euro S\&P submission, we expect more useful content than what is in this paper, so you have 13 pages of body text.
+
 \bibliographystyle{plain}

### 2023-11-03 paper/main.tex
tweaks to the test section

--- 2023-11-03 paper/main.tex
@@
 
-\section{Statistical testing for MPC insecurity}
+\section{\toolname: a Statistical Test for Security}
 
-\subsection{Test algorithm}
+This section describes \toolname, our approach for using statistical independence testing to find security bugs in MPC protocols.
+The approach is based on the \emph{fast conditional independence test} proposed by Chalupka et al.~\cite{chalupka2018fast}.
+
+\subsection{Conditional Independence Testing by Classification}
+
+Chalupka et al.~\cite{chalupka2018fast} present a high-dimensional conditional independence test based on the observation that if $P(X | Y, Z) = P(X | Y)$ (i.e.
+$X$ and $Z$ are conditionally independent given $Y$), then $Y$ and $Z$ together will not be more useful for predicting $X$ than $Y$ alone.
+They leverage this observation to design a test based on training machine learning classifiers: one classifier predicts $X$ from both $Y$ and $Z$, while the other predicts $X$ from $Y$ alone.
+If the first classifier yields better test error than the second, then the conditional independence property likely does not hold.
+
+\subsection{The \toolname Algorithm}
 
@@
 
-We use the techniques from \cite{chalupka2018fast}
+We use the techniques from Chalupka et al.~\cite{chalupka2018fast}
 to directly test for conditional dependence without any static analysis of the subject protocol.
@@
 
-\subsection{Theoretical issues}
+\subsection{Interpreting Test Results}
 
@@
 The usual trade-offs of significance testing apply.
-Setting $\alpha$ too low will cause programs who's insecure-ness is only marginally discernible to the D-Trees
+Setting $\alpha$ too low will cause programs whose insecure-ness is only marginally discernible to the D-Trees
 to pass the test.
@@
 and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.
-
-TODO: link to github or other archive?
+Our implementation and the code to reproduce our experimental evaluation are available as open-source.\footnote{(Link removed for anonymous submission)}
 

### 2023-11-03 paper/main.tex
add a couple of comments

@@
 
+\todo{we should maybe introduce decision trees?}
+
 \subsection{The \toolname Algorithm}
@@
 \langname supports macros, but they are extremely limited.
-Finally, \langname supports 1-of-N oblivious transfer as a primitive operation.
-This allows us to test (correct and incorrect) implementations of the classic GMW protocol.
 
-\paragraph{Oblivious Transfer.}
+\langname supports 1-of-N oblivious transfer as a primitive operation.
 Oblivious transfer~\cite{todo} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
@@
 %
-Oblivious transfer is a building block for many other protocols.
+Oblivious transfer is a building block for many other protocols, including the GMW protocol described later.
 
-TODO: refer to an example figure and/or an appendix.
+\todo{refer to an example figure and/or an appendix.}
 

### 2023-11-03 paper/main.tex

--- 2023-11-03 paper/main.tex
@@
 the techniques used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.
-
-\section{Related Work}
-
-\subsection{Formal-methods for cryptography}
-Although a useful language-driven test for MPC security remains out of reach,
-a variety of useful work aproaches the subject from different directions.
-
-Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
-use EasyCrypt to manually show the security of GMW.
-Their definitions generalize to protocols with efficiently invertible functionalities.
-They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
-If their system could be automated and paired with language constraints to ensure invertibility,
-then it could apply to many, but certainly not all, use cases.
-
-Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
-give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
-Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.
-
-Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
-give a powerful language system that can check if one protocol is a valid simulator of another,
-and show that it works on existing MPC protocols like GMW.
-The user of the system needs to write a second protocol
-representing the composition of the ideal-world functionality and an explicit $\mathtt{SIM}$,
-and then build a high-level proof of the equivalence in the provided equational logic.
-
-The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
-gives fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
-Their system assumes (at least one) secure implementation of MPC on circuits.
-
-$\lambda_\textbf{obliv}$~\cite{darais2019language} %
-gives a type system that ensures uniform and independent randomness of key values throughout a program.
-Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
-but this would of course prevent any non-vacuous functionality from being implemented.
-
-\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
-gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
-They can type-check convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
-there's no way to verify an implementation of MPC.
-
-Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
-also use a type system to ensure safe usage of cryptographic primitives,
-but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
-Like OWL, their type system assumes a secure implementation of the encryption scheme.
-
-\subsection{PPLs for Inference}
-
-We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
-A SPN is a directed-graph representation of a multi-variate distribution.
-They are compact, allow efficient inference, and can represent values symbolically as needed,
-so testing for conditional independence is practical.
-However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
-because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.
-
-The \textsc{Dice} language~\cite{holtzen2020scaling} translates programs into binary decision diagrams (BDDs),
-analogous to the way we tried to translate them into SPNs.
-BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
-but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
-a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
-
-Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
-is a program logic for reasoning about probability distributions and PPLs.
-It more expressive than the original probabilistic separation logic~\cite{barthe2019probabilistic},
-but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.
-
-\subsection{Probabilistic testing}
-
-Ding et al.~\cite{ding2018detecting} % https://arxiv.org/pdf/1805.10277.pdf
-and more recently Zhang et al.~\cite{zhang2020testing} % https://dl.acm.org/doi/10.1145/3428233
-use large samples of program executions to test for violations of differential privacy,
-showing that such techniques can work on probabilistic hyper-properties.
-Since differential privacy is only concerned with program inputs and outputs,
-the dimensionality of the problem is somewhat smaller than MPC.
-
-Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
-who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
-Specifically, if decision trees (D-Trees) trained on random variables $A$ and $C$
-do better at predicting the value of a random variable $B$ than D-Trees trained only on $C$,
-then $A$ and $B$ are not independent conditioned on $C$.
-The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
-In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
-but in our adaptation we use fresh program traces for everything.
 
@@
 
-Our \toolname python script consumes a stream of fixed-length bit vectors as a CSV.
+Our \toolname Python script consumes a stream of fixed-length bit vectors as a CSV.
 Annotation of which columns are available in the real and ideal worlds is provided in the header.
 The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
-The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
+The individual ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
 and they're trained in parallel with \texttt{joblib.Parallel}.
@@
 
+\section{Related Work}
+
+\subsection{Formal-methods for cryptography}
+Although automated language-based verification of MPC security is an open question,
+a variety of useful work approaches the subject from different directions.
+
+Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
+use EasyCrypt to manually show the security of GMW.
+Their definitions generalize to protocols with efficiently invertible functionalities.
+They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
+Their approach requires manual proof in EasyCrypt.
+
+Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
+give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
+However, the framework is not amenable to automation.
+
+Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
+give a powerful language system that can check if one protocol is a valid simulator of another,
+and show that it works on existing MPC protocols like GMW.
+The user of the system needs to write a second protocol
+representing the composition of the ideal-world functionality and an explicit $\mathtt{SIM}$,
+and then build a high-level proof of the equivalence in the provided equational logic.
+
+The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
+gives fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
+Their system assumes (at least one) secure implementation of MPC on circuits.
+
+$\lambda_\textbf{obliv}$~\cite{darais2019language} %
+gives a type system that ensures uniform and independent randomness of key values throughout a program.
+Their system is automated, and could be adapted to ensure perfect noninterference between honest secrets and corrupt views,
+but does not support the kind of \emph{conditional} independence required for most MPC protocols.
+
+\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
+gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
+OWL is designed for secure protocols that reveal nothing to the adversary, and does not support verification of MPC security.
+
+Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
+also use a type system to ensure safe usage of cryptographic primitives,
+but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
+Like \textsc{Owl}, their type system assumes a secure implementation of the encryption scheme and does not support verification of MPC security.
+
+\subsection{PPLs for Inference}
+
+Several PPLs directly support efficient inference and some forms of independence testing.
+A Sum Product Network (SPN)~\cite{poon2011sum} is a directed-graph representation of a multi-variate distribution;
+SPPL~\cite{saad2021sppl} is a PPL that performs inference using SPNs.
+SPNs are compact, allow efficient inference, and can represent values symbolically as needed,
+so testing for conditional independence is practical.
+However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
+because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.
+
+The \textsc{Dice} language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
+BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
+but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
+a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.
+
+\textsc{Lilac}~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
+is a program logic for reasoning about probability distributions and PPLs.
+It more expressive than the original probabilistic separation logic,
+but does not support automation.
+
+\subsection{Probabilistic testing}
+
+Ding et al.~\cite{ding2018detecting} % https://arxiv.org/pdf/1805.10277.pdf
+and more recently Zhang et al.~\cite{zhang2020testing} % https://dl.acm.org/doi/10.1145/3428233
+use large samples of program executions to test for violations of differential privacy,
+showing that such techniques can work on probabilistic hyper-properties.
+Since differential privacy is only concerned with program inputs and outputs,
+the dimensionality of the problem is somewhat smaller than MPC.
+
+Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
+who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
+Specifically, if decision trees (D-Trees) trained on random variables $A$ and $C$
+do better at predicting the value of a random variable $B$ than D-Trees trained only on $C$,
+then $A$ and $B$ are not independent conditioned on $C$.
+The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
+In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
+but in our adaptation we use fresh program traces for everything.
+
 \section{Future Work}

### 2023-11-03 paper/main.tex

@@
 
-\section{Statistical testing for MPC insecurity}
+\section{\toolname: a Statistical Test for Security}
 
-\subsection{Test algorithm}
+This section describes \toolname, our approach for using statistical independence testing to find security bugs in MPC protocols.
+The approach is based on the \emph{fast conditional independence test} proposed by Chalupka et al.~\cite{chalupka2018fast}.
+
+\subsection{Conditional Independence Testing by Classification}
+
+Chalupka et al.~\cite{chalupka2018fast} present a high-dimensional conditional independence test based on the observation that if $P(X | Y, Z) = P(X | Y)$ (i.e.
+$X$ and $Z$ are conditionally independent given $Y$), then $Y$ and $Z$ together will not be more useful for predicting $X$ than $Y$ alone.
+They leverage this observation to design a test based on training machine learning classifiers: one classifier predicts $X$ from both $Y$ and $Z$, while the other predicts $X$ from $Y$ alone.
+If the first classifier yields better test error than the second, then the conditional independence property likely does not hold.
+
+\todo{we should maybe introduce decision trees?}
+
+\subsection{The \toolname Algorithm}
 
@@
 
-We use the techniques from \cite{chalupka2018fast} to build \toolname,
+We use the techniques from Chalupka et al.~\cite{chalupka2018fast} to build \toolname,
 which directly tests for the above conditional dependence without any static analysis of the subject protocol.
@@
 
-See Section~\ref{sec:public-repo} for more details.
+Our implementation and the code to reproduce our experimental evaluation are available as open-source.\footnote{(Link removed for anonymous submission)}
 
@@
 
-\section*{Data Availability}\label{sec:public-repo}
-
-TODO.
-
-This section, and everything following, does not count within the body text page limit requirement (so this paper would satisfy a page limit of 1).
-For your Euro S\&P submission, we expect more useful content than what is in this paper, so you have 13 pages of body text.
-
 \bibliographystyle{plain}

### 2023-11-03 paper/main.tex
explain what a simulator is
describe decision trees

@@
 
-\todo{this is the first place the simulator is mentioned, but it's not described}
 \begin{definition}[Passive MPC Security]\label{def:passive-mpc}
@@
 anything the real-world adversary can compute the ideal-world adversary can also compute!
+The hypothetical function $\mathtt{SIM}$ is called a "simulator";
+it simulates in the ideal world what the real-world adversary would see.
+Often MPC security proofs will give an explicit simulator function.
 
@@
 
-\todo{we should maybe introduce decision trees?}
+Chalupka et al. use decision trees (D-Trees) as their ML classifiers.
+Because the process requires training many copies of the model, a light-weight model like a D-Tree is preferable;
+we also use D-Trees for \toolname.
 

### 2023-11-03 paper/main.tex
describe decision trees

@@
 we also use D-Trees for \toolname.
+In principal, \textit{any} ML classifier would be ok to use; the choice is partly an engineering question
+and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
 
@@
 To actually test a protocol, \toolname uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
-One of the two models has access to the real-world views, and the only has access to the ideal-world views.
-In principal, \textit{any} ML classifier is ok to use; the choice is partly an engineering question
-and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
+One of the two models has access to the real-world views, and the other only has access to the ideal-world views.
 Generating the data for these tests requires running the protocol many times.

### 2023-11-03 paper/main.tex
small tweaks to evaluation

@@
 
-\subsection{Random generation of programs}
+\subsection{Bug-Finding in Generated Protocols}
 \label{sec:random-protocols}
 
-The goal of this work is a tool that can detect mistakes in MPC protocols.
-Such a tools is only useful if there if new protocols and implementations are being written.
-Therefore, while it is critical to demonstrate that our tool can detect errors manually edited into known-secure protocols,
-we \textit{also} need to show that the tool can detect a lack of security in entirely novel and maximally diverse protocols.
-We attempt to show this by randomly generating hundreds [can we say thousands?] of protocols and testing them with different test parameters.
+The goal of this work is a tool that can detect mistakes in diverse MPC protocols.
+% Such a tools is only useful if there if new protocols and implementations are being written.
+% Therefore, while it is critical to demonstrate that our tool can detect errors manually edited into known-secure protocols,
+% we \textit{also} need to show that the tool can detect a lack of security in entirely novel and maximally diverse protocols.
+We evaluate the generality of the approach by randomly generating hundreds of protocols and testing them with different test parameters.
 We also use a batch of randomly generated protocols to asses the effects of the three test parameters
@@
 
-The probability that a randomly generated protocol will be MPC secure is not completely negligible.
+Randomly-generated protocols are likely to be insecure, but the probability of generating a secure protocol is not zero.
 In particular, some generator settings will be biased toward vacuous protocols in which nothing is communicated
 or vacuously secure protocols in which the honest secrets are fully revealed by the ideal functionality.
-(Never mind that such protocols are useless; none of the protocols generated in this experiment are \textit{useful}.)
-That said, for reasonable generator settings we find that very few generated protocols
+% (Never mind that such protocols are useless; none of the protocols generated in this experiment are \textit{useful}.)
+For reasonable generator settings we find that very few generated protocols
 continue to appear secure as we increase the power of the test.
@@
 
-The results of these experiments suggest that \toolname scales linearly with protocol size and test power, and is capable of detecting bugs in real-world protocols.
+The results of these experiments suggest that \toolname scales linearly with protocol size and test power, and is capable of detecting bugs in both randomly-generated and real-world protocols.
 The test typically runs in just seconds to minutes.
@@
 
+This Appendix contains additional experimental results excluded from the main body of the paper:
+%
+\begin{itemize}
+\item Figure~\ref{fig:extra1} presents additional results for $n$-bit addition.
+\item Figure~\ref{fig:extra2} presents additional results for $n$-bit less-than comparison.
+\item Figure~\ref{fig:extra3} presents additional results for $n$-bit Beaver triple generation.
+\end{itemize}
+
 \input{full_figures.tex}

### 2023-11-03 paper/main.tex
add OT ref

@@
 \langname supports 1-of-N oblivious transfer as a primitive operation.
-Oblivious transfer~\cite{todo} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
+Oblivious transfer~\cite{kilian1988founding} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
 In 1-out-of-2 OT, $S$ provides 2 secret inputs $x_0$ and $x_1$ and $R$ provides a secret \emph{selection bit} $b$.

### 2023-11-03 paper/main.tex
example cho

@@
 
-\todo{refer to an example figure and/or an appendix.}
+An example \langname program is provided in Appendix~\ref{sec:example-cho}.
 
@@
 
+\section{Example \langname program}\label{sec:example-cho}
+
+The following \langname uses oblivious-transfer based GMW to evaluate a circuit representing a "less than" function
+between the two parties' inputs, interpreted as 2-bit binary numbers.
+It is slightly shorter than the ones we used for our experiments;
+to provide the closest possible comparison between different implementations in our experiments,
+the \langname protocols we used included some "no-op" statements.
+
+\begin{small}
+\begin{verbatim}
+-- two_bit_less_than.cho
+MACRO secret_share(P1(x), P2()) AS
+  s1 = FLIP @P1
+  s2 = x + s1
+  SEND s2 TO P2
+ENDMACRO
+MACRO and_gmw(P1(x2, y2), P2(x1, y1)) AS
+  out2 = FLIP @P2
+  g1_s2_00 = out2 + ((x1 + 0) ^ (y1 + 0))
+  g1_s2_01 = out2 + ((x1 + 0) ^ (y1 + 1))
+  g1_s2_10 = out2 + ((x1 + 1) ^ (y1 + 0))
+  g1_s2_11 = out2 + ((x1 + 1) ^ (y1 + 1))
+  out1 = OBLIVIOUSLY [[g1_s2_00,
+                       g1_s2_01]?y2,
+                      [g1_s2_10,
+                       g1_s2_11]?y2
+                     ]?x2 FOR P1
+ENDMACRO
+MACRO reveal(P1(x1), P2(x2)) AS
+  SEND x1 TO P2
+  SEND x2 TO P1
+  y = x1 + x2
+ENDMACRO
+-- Read secrets
+x0 = SECRET @P1
+x1 = SECRET @P1
+y2 = SECRET @P2
+y3 = SECRET @P2
+-- Set up shares
+DO secret_share(P1(x0), P2()) GET(x0_1=s1,
+                                  x0_2=s2)
+DO secret_share(P1(x1), P2()) GET(x1_1=s1,
+                                  x1_2=s2)
+DO secret_share(P2(y2), P1()) GET(y2_1=s2,
+                                  y2_2=s1)
+DO secret_share(P2(y3), P1()) GET(y3_1=s2,
+                                  y3_2=s1)
+-- Circuit evaluation
+g3_1 = ~y2_1
+g3_2 = y2_2
+DO and_gmw(P1(x0_1, g3_1), P2(x0_2, g3_2)
+          ) GET(g4_1=out1, g4_2=out2)
+g5_1 = ~x0_1
+g5_2 = x0_2
+DO and_gmw(P1(y2_1, g5_1), P2(y2_2, g5_2)
+          ) GET(g6_1=out1, g6_2=out2)
+g7_1 = ~y3_1
+g7_2 = y3_2
+g8_1 = ~g6_1
+g8_2 = g6_2
+DO and_gmw(P1(g7_1, g8_1), P2(g7_2, g8_2)
+          ) GET(g9_1=out1, g9_2=out2)
+DO and_gmw(P1(x1_1, g9_1), P2(x1_2, g9_2)
+          ) GET(g10_1=out1, g10_2=out2)
+DO and_gmw(P1(g4_1, g10_1), P2(g4_2, g10_2)
+          ) GET(g11_1=out1, g11_2=out2)
+g12_1 = g4_1 + g10_1
+g12_2 = g4_2 + g10_2
+g13_1 = g11_1 + g12_1
+g13_2 = g11_2 + g12_2
+-- Reveal output
+DO reveal(P1(g13_1), P2(g13_2)) GET(r0=y)
+OUTPUT r0
+\end{verbatim}
+\end{small}
+
 \end{document}

### 2023-11-03 paper/main.tex
line edits and removed commented sections

@@
     \hline
-    %Oblivious Transfer & Building block & TODO \\
-%    OT Extension & Building block & TODO \\
     GMW~\cite{goldreich2019play, goldreich2009foundations} & Oblivious Transfer\\
@@
     \hline
-    \textbf{Instantiated Functions} & \textbf{Leakage}\\
-    \hline
-    $n$-bit Addition & All \\
+    \textbf{Instantiated Functions} & \textbf{Invertibility}\\
+    \hline
+    $n$-bit Addition & Complete \\
     $n$-bit Less-than Comparison & Partial \\
@@
 The third, an $n$-bit Beaver triple generator, leaks no information about the honest party's input, since its output to each party is a single secret share of the Beaver triple.
-
-% \paragraph{$n$-bit Addition.}
-% The simplest protocol in our evaluation is a basic summation protocol, which adds up the inputs from the two participating parties.
-To accomplish this task, the protocol essentially simulates a hardware adder circuit---for each bit in the binary representation of the two input numbers, the protocol performs bitwise operations to compute the sum and carry bit.
-To ensure security, the input values are \emph{secret shared} using additive secret sharing; to secret share a bit, a party selects a random bit to represent one share, and computes the other share by XORing the random bit with the secret bit.
-In \langname, $P_1$ can secret share the bit $x$ with $P_2$ as follows:
-% %
-% \begin{verbatim}
-%   s1 = FLIP @P1
-%   s2 = x + s1
-%   SEND s1 TO P2
-% \end{verbatim}
-% %
-% $P_1$ can safely send one share to $P_2$, as long as $P_1$ keeps the other share for themselves---neither share alone reveals anything about the secret bit.
-To add (i.e.
-XOR) two secret-shared bits, each party can simply add its shares for the two bits to obtain one share of the result.
-To multiply (i.e.
-AND) two secret-shared bits, the parties must use oblivious transfer (described next).
-
-% Our $n$-bit addition protocol generates a \langname source file for a given bitwidth.
-In our evaluation, we test several different bitwidths to examine the effect of the size of each party's view on the scalability and bug-finding capability of \toolname.
 
@@
 
-% \paragraph{Oblivious Transfer Extension.}
-% Protocols like GMW (described later) require the use of OT for each AND gate---for large circuits, this means thousands or millions of OTs.
-\emph{OT extension}~\cite{todo} is an approach for expanding $k$ ``base'' OTs into $m \gg k$ OTs.
-OT extension can significantly improve the performance of protocols that use many OTs.
-
-% Our empirical evaluation demonstrates \toolname's ability to detect bugs in secure protocols.
-% To answer RQ1, we generate \emph{random protocols} of varying sizes, and evaluate \toolname's ability to show they are insecure.
-The details appear in Section~\ref{sec:experiment-1:-random}.
-To answer RQ2, we evaluate \toolname on the case study protocols described in Section~\ref{sec:case-study-protocols}.
-The details appear in Section~\ref{sec:experiment-2:-real}.
-In both cases, \toolname is able to detect insecurity in the target protocols in under a minute.
-
 \begin{figure*}
@@
 
-We run \toolname on the unmodified protocol in order to establish the protocol's security.
-Then, we mutate the protocol in several ways to introduce security bugs, and test \toolname's ability to detect these bugs.
-Specifically, our mutations are:
+We run \toolname on the unmodified protocol in order to establish the protocol's security
+and show that \toolname doesn't flag secure protocols as insecure.
+Then, we mutate the protocol in several ways to introduce security bugs, and test \toolname's ability to detect these bugs. Specifically, our mutations are:
 %
@@
 \item \textbf{Biased AND}: uses biased randomness as part of the evaluation of multiplication gates.
-\item \textbf{Accidental secret}: accidentally sends a secret held by the honest party  to the corrupt party.
-\item \textbf{Accidental gate}: accidentally sends an honest party's result from the evaluation of a multiplication gate to the corrupt party.
+\item \textbf{Accidental secret}: sometimes, at random, accidentally sends a secret held by the honest party  to the corrupt party.
+\item \textbf{Accidental gate}: sometimes, at random, accidentally sends an honest party's result from the evaluation of a multiplication gate to the corrupt party.
 \end{itemize}
@@
 
-\toolname detects \textbf{accidental gate} bugs only in the $n$-bit less-than comparison based on the GMW framework.
-In the Beaver-triple framework, we hypothesize that the decision trees are not capable of the ``backwards computation'' required to leverage the leaked information; in the other cases, the function's design means that AND gate outputs do not provide additional leakage beyond the ideal functionality.
-
-\toolname fails to detect \textbf{biased AND} and \textbf{accidental gate} bugs in the case of $n$-bit Beaver triple generation because the result of each AND gate is also one of the protocol's outputs---so a bug which reveals it does not leak any new information.
+\toolname detects \textbf{accidental gate} bugs in the $n$-bit less-than comparison only in the OT-based GMW framework.
+In the Beaver-triple framework, we hypothesize that the decision trees are not capable of the ``backwards computation''
+required to leverage the leaked information
+
+In other cases, the function's design means that AND gate outputs do not provide additional leakage beyond the ideal functionality,
+and therefore do not compromise security.
+\toolname ignores \textbf{biased AND} and \textbf{accidental gate} bugs in the case of $n$-bit Beaver triple generation
+because the result of each AND gate is also one of the protocol's outputs---so a bug which reveals it does not leak any new information.
 \toolname detects \textbf{accidental gate} bugs for the $n$-bit Beaver triple generation function when the Beaver-triple-based protocol is used, because biasing the randomness for AND gates reveals information about the Beaver triples themselves, from which it is possible to deduce the honest party's shares of the inputs.

### 2023-11-03 paper/main.tex
line edits

@@
   It successfully detects several different mistakes and biases introduced into two different implementations of the classic GMW protocol.
-  Applied to thousands of randomly generated protocols, it identifies nearly all of them as insecure.
+  Applied to hundreds of randomly generated protocols, it identifies nearly all of them as insecure.
   We also include an analysis of the parameters of the test, and discussion of what makes detection of MPC (in)security difficult.
@@
 \toolname works by defining security in terms of conditional independence, then performing a statistical test which attempts to falsify this conditional independence property.
-In MPC protocols, the inputs of the honest parties should be independent of the \emph{views} (i.e.
-the observed message values) of the corrupt parties, conditioned on the protocol's output.
+In MPC protocols, the inputs of the honest parties should be independent of the \emph{views} (\ie{} the observed message values) of the corrupt parties, conditioned on the protocol's output.
 Conditioning on the protocol's output is vital for the security definition, since it allows the protocol to \emph{intentionally} leak some information about the inputs via the output---as nearly all MPC protocols do.
@@
 This formulation of security as conditional independence puts MPC security out of reach of all automated verification tools (and many manual ones).
-While automated tools have been used to verify security in cryptographic protocols~\cite{gancher2023owl, barthe2019probabilistic, darais2019language, fournet2011information}  (e.g.
-key exchange protocols), these tools are limited to protocols whose \emph{outputs reveal nothing} to the adversary.
+While automated tools have been used to verify security in cryptographic protocols~\cite{gancher2023owl, barthe2019probabilistic, darais2019language, fournet2011information}  (\eg{} key exchange protocols), these tools are limited to protocols whose \emph{outputs reveal nothing} to the adversary.
 They are capable of showing independence, but not the kind of \emph{conditional} independence required for MPC security.
@@
 
-\toolname is designed to fill the same role as tools like \textsc{QuickCheck}: it can be used to quickly and automatically detect many classes of bugs in MPC protocol implementations, without requiring the programmer to spend time building a manual security proof.
+\toolname is designed to fill the same role as tools like \textsc{QuickCheck}: it can be used to quickly and automatically detect many classes of bugs in MPC protocol implementations, without investing the user's time in a manual security proof.
 Using \toolname requires only that key runtime values can be captured during program execution;
@@
 to compute a function that depends on all of their private inputs,
-without any of them learning anything except the final result.
+without any of them learning anything except the final result\footnote{
+    It is possible for each party to have a distinct final result.
+}.
 Implementations that work on arbitrary functions (representable as circuit computations in some field)
@@
 
-Since MPC security can be framed as a test of conditional independence, \termOfArt{probabilistic programming languages} (PPLs) with efficient inference and independence tests could be used to verify MPC security.
+Since MPC security can be framed as a test of conditional independence
+(Section~\ref{sec:algorithm}),
+\termOfArt{probabilistic programming languages} (PPLs) with efficient inference and independence tests could hypothetically be used to verify MPC security.
 %
@@
 }, and
-(2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
+(2) The secret inputs of the honest parties are not stochastic but the adversary can be asked to provide a distribution over
 possible secrets representing their prior belief so that the secret inputs can also be modeled as random variables.
@@
 
-\subsection{The \toolname Algorithm}
+\subsection{The \toolname Algorithm}\label{sec:algorithm}
 
@@
 \langname supports 1-of-N oblivious transfer as a primitive operation.
-Oblivious transfer~\cite{kilian1988founding} (OT) is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
+Oblivious transfer (OT)~\cite{kilian1988founding} is a 2-party protocol that allows one party ($R$, the \emph{receiver}) to select one out of several secret inputs provided by the other party ($S$, the \emph{sender}).
 In 1-out-of-2 OT, $S$ provides 2 secret inputs $x_0$ and $x_1$ and $R$ provides a secret \emph{selection bit} $b$.
@@
 \begin{figure*}
-  \label{fig:sankey}
   \centering
@@
 \caption{Experimental results: performance on random protocols under increasing test power.}
+\label{fig:sankey}
 \end{figure*}
@@
 \begin{figure}
-  \label{fig:linear-time}
   \centering
@@
 \caption{Experimental results: runtime with increasing test power.}
+\label{fig:linear-time}
 \end{figure}
@@
 \begin{figure*}
-  \label{fig:cube-search}
   \centering
@@
 \caption{Experimental results: trade-offs in power and runtime under different test parameters.}
+\label{fig:cube-search}
 \end{figure*}
@@
 Each 3D heatmap shows aggregate data for 200 randomly generated \toolname files.
-Because is mostly driven by the amount of data consumed ($\texttt{iters} \cdot (\texttt{trainN} + \texttt{testN})$),
+Because runtime is mostly driven by the amount of data consumed ($\texttt{iters} \cdot (\texttt{trainN} + \texttt{testN})$),
 all three affect runtime linearly.
@@
 and show that \toolname doesn't flag secure protocols as insecure.
-Then, we mutate the protocol in several ways to introduce security bugs, and test \toolname's ability to detect these bugs. Specifically, our mutations are:
+Then we mutate the protocol in several ways to introduce security bugs and test \toolname's ability to detect these bugs. Specifically, our mutations are:
 %
@@
 For larger bitwidths, the decision trees are no longer able to perform this reconstruction, and the protocols are classified as insecure.
-The other functions---which are not vacuously secure---do not display this behavior.
 
@@
 
-\toolname detects \textbf{accidental gate} bugs in the $n$-bit less-than comparison only in the OT-based GMW framework.
+\toolname detects \textbf{accidental gate} bugs in the $n$-bit-less-than comparison only in the OT-based framework.
 In the Beaver-triple framework, we hypothesize that the decision trees are not capable of the ``backwards computation''
@@
 Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
-who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
+who show that batches of ML classifiers can be used to check conditional independence.
 Specifically, if decision trees (D-Trees) trained on random variables $A$ and $C$
@@
 
-Some further works is needed to turn \toolname into a deploy-able tool;
+Some further work is needed to turn \toolname into a deploy-able tool;
 our experiments so far are not sufficient to calculate sensible default settings,
@@
 Decision trees are not the only candidate for the underlying ML model, and our evaluation suggests that they are incapable of performing some of the ``backwards computation'' needed to demonstrate insecurity;
-a more capable model (\eg{} a neural network) may be more capable of detecting these kinds of security bugs.
+a deeper model (\eg{} a neural network) may be more capable of detecting these kinds of security bugs.
 Other cryptographic properties besides MPC can likely be framed in terms of conditional independence;

### 2024-03-11 paper/main.tex
add funding acknowledgments

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{natbib}
+%\usepackage{natbib}
 
@@
     {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}}~}
-%\newcommand{\mynote}[2]{}
 \newcommand{\todo}[1]{\mynote{TODO}{#1}}
@@
 
-%\author{First1 Last1}
-%%\authornote{with author1 note}        %% and a great more deleted...
+\author{\IEEEauthorblockN{Mako Bates}
+\IEEEauthorblockA{\textit{University of Vermont} \\
+Burlington, Vermont, US \\
+mako.bates@uvm.edu
+}
+\and
+\IEEEauthorblockN{Joseph P. Near}
+\IEEEauthorblockA{\textit{University of Vermont} \\
+Burlington, Vermont, US \\
+jnear@uvm.edu
+}
+}
 
@@
 \end{itemize}
-
-% In this work we present and evaluate a statistical test of the security of MPC protocols.
-% First we show how MPC security relates to inference and equivalence of probabilistic distributions,
-% and then we demonstrate that existing tests of distributional equivalence can, with minor modification, be used to refute MPC security.
-% Our tool is adaptable to the wide range of cryptographic security properties that can be defined in terms of distributional equivalence,
-% and our tool is applicable to any software system that
-% A) can be run 10k-1M times without great inconvenience, and
-% B) allows certain intermediate values (of which consumed randomness is the least convenient) to be captured for analysis.
 
@@
 The goal of this work is a tool that can detect mistakes in diverse MPC protocols.
-% Such a tools is only useful if there if new protocols and implementations are being written.
-% Therefore, while it is critical to demonstrate that our tool can detect errors manually edited into known-secure protocols,
-% we \textit{also} need to show that the tool can detect a lack of security in entirely novel and maximally diverse protocols.
 We evaluate the generality of the approach by randomly generating hundreds of protocols and testing them with different test parameters.
@@
 
+\section*{Acknowledgments}
+
+This material is based upon work supported by the National Science Foundation under Grant No. 2238442 and by the Cold Regions Research and Engineering Laboratory (ERDC-CRREL)
+under Contract No.
+W913E521C0003.
+Any opinions, findings and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation or the Cold Regions Research and Engineering Laboratory.
+
 \bibliographystyle{plain}
